---
title: "BCFtools Stats Report"
author:
    - "Created using HARPY"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    highlight: tango
    number_sections: no
    theme: default
    df_print: kable
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
---
```{r load environment, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(magrittr)
library(tidyr)
library(DT)
library(plotly)
library(RColorBrewer)
#dataL <- readLines("~/checkstats.txt")
dataL <- readLines(snakemake@input[[1]])
bcf <- gsub(".stats$", ".bcf", snakemake@input[[1]])
```
### File: `r bcf`

## General Information
This reflects the general information stored in the records of `r bcf`.
```{r General Stats, echo = FALSE, message = FALSE, warning = FALSE}
.snL <- grepl("^SN", dataL)
sn <- read.table(text=dataL[.snL], sep = "\t")[,3:4]
names(sn) <- c("Metric", "Number")
sn$Metric <- gsub("number of ", "", sn$Metric)
rownames(sn) <- sn$Metric
sn[2]
```


## Individual Statistics
These statistics correspond to the `r sn$Number[1]` samples in `r bcf`.
```{r Individual Stats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
.pscL <- grepl("^PSC", dataL)
psc <- read.table(text=dataL[.pscL])[ ,3:14]
names(psc) <- c("Sample", "HomozygousRef", "HomozygousAtl", "Heterozygotes", "Transitions", "Transversions", "Indels",	"MeanDepth", "Singletons",	"HapRef", "HapAlt", "Missing")
psc$Homozygotes <- psc$HomozygousRef + psc$HomozygousAtl
tidy_psc <- pivot_longer(psc[,c(1,4,5,6,7,9,12,13)], -Sample , names_to = "Metric", values_to = "Count")
psc <- psc[,c(1,8,12,2,3,13,4,5,6,7,9,10,11)]

pscplot <- ggplot(data = tidy_psc, mapping = aes(x = Count, y = Sample, color = Metric)) +
  geom_point(size = 2) +
  labs(title = "Individual Statistics") +
  theme_bw() +
  scale_x_continuous(n.breaks=9) +
  xlab("Count") +
  scale_color_brewer(palette = "Dark2") +
  theme(panel.grid.minor = element_blank(), panel.grid.major.y = element_blank())

ggplotly(pscplot)
DT::datatable(psc, rownames = F, filter = "top", extensions = 'Buttons', options = list(dom = 'Brtip', buttons = c('csv', 'pdf'), scrollX = TRUE))
```

## Locus Statistics
These are per-locus statistics, which correspond to the Depth (`DP`) calculations of `bcftools stats`
```{r Locus Stats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
.dpL <- grepl("^DP", dataL)
if(length(.dpL)) < 1{
  cat(paste0("No DP section in", snakemake@input[[1]], " found\n")
} else {
  dp <- read.table(text=dataL[.dpL])[ ,3:7]
  names(dp) <- c("Bin", "Genotypes", "PercentGenotypes", "NumberSites", "PercentSites")
  dpplot <- ggplot(data = dp, mapping = aes(x = Bin, y = NumberSites)) +
    geom_bar(stat = "identity", color = "indianred", fill = "indianred") +
    scale_y_continuous(n.breaks=10) +
    labs(title = "Depth per Site") +
    ylab("Number of Sites") +
    theme_minimal() +
    theme(panel.grid.minor.y = element_blank())

  ggplotly(dpplot)
  DT::datatable(dp, rownames = F, filter = "top", extensions = 'Buttons', options = list(dom = 'Brtip', buttons = c('csv', 'pdf')))
}
```

## Insertions and Deletions
Below is a distribution of insertions and deletions based on length and frequency.
```{r Indel Stats, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
.iddL <- grepl("^IDD", dataL)
idd <- read.table(text=dataL[.iddL])[ ,3:4]
names(idd) <- c("Length", "Count")
idd$Type <- idd$Length > 0
idd$Type <-  gsub(TRUE, "Insertion", idd$Type)
idd$Type <-  gsub(FALSE, "Deletion", idd$Type)
iddplot <- ggplot(data = idd, mapping = aes(x = Length, y = Count,  color = Type)) +
  geom_point() +
  scale_color_manual(values=c("indianred","#56B4E9"))+
  labs(title = "Insertion-Deletion Distribution") +
  xlab("indel length") +
  ylab("number of sites") +
  theme_minimal()

ggplotly(iddplot)
```

