---
title: "Leviathan Variant Calling Summary"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    vertical_layout: scroll
    favicon: "https://raw.githubusercontent.com/pdimens/harpy/docs/static/favicon_dark.png"
---

```{r load environment, echo=FALSE, message=FALSE, warning=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(magrittr)
library(DT)
library(plotly)
library(circlize)
library(ComplexHeatmap)
```
```{css zoom-lib-src, echo = FALSE, message = FALSE, warning = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.7.1/jquery.min.js"
```
```{js zoom-jquery, echo = FALSE, message = FALSE, warning = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

# General Stats
## Intro
### introtext {.no-title}
```{r}
statsfile <- snakemake@input[["statsfile"]]
samplename <- gsub(".stats", "", basename(statsfile))
```

<h1> Structural Variants: `r samplename`</h1>
Below is a series of tables and plots detailing the structural variants
identified by [LEVIATHAN](https://github.com/morispi/LEVIATHAN) [(preprint)](https://doi.org/10.1101/2021.03.25.437002). The variants are given by:

```{r table2, echo=FALSE, message=FALSE, warnings=FALSE}
knitr::kable(
  data.frame(
  "Variant" = c("INV", "DUP", "DEL","BND"),
  "Name" = c("Inversion", "Duplication", "Deletion", "Breakend"),
  "Description" = c(
    "a segment that broke off and reattached within the same chromosome, but in reverse orientation",
    "a type of mutation that involves the production of one or more copies of a gene or region of a chromosome",
    "a type of mutation that involves the loss of one or more nucleotides from a segment of DNA",
    "the endpoint of a structural variant, which can be useful to explain complex variants"
    )
  )
)
```

This tab (`General Stats`) shows overview information. Clicking `Per-Contig Plots`
in the navigation bar at the top of this page will show you interactive
plots detailing all variants identified.

```{r}
statsfile <- snakemake@input[["statsfile"]]
#statsfile <- "/home/pdimens/test.stats"
sv <- read.table(statsfile, header = T,
  colClasses = c("factor", "factor", "integer", "integer", "character", "factor", "integer", "integer")
)
if (nrow(sv) == 0) {
  cat(paste0("There are no variants in the file ", "`", bcf, "`"))
  knitr::knit_exit()
}
```
```{r contigSizes, echo = FALSE}
bcf <- snakemake@input[["bcf"]]
#bcf <- "/home/pdimens/test.bcf"
hd <- system2("bcftools", args =  paste("head", bcf), stdout = T)
fa.sizes <- hd[grep("contig=", hd)]
fa.sizes <- as.data.frame(do.call(rbind, strsplit(fa.sizes, ",length="))) %>% arrange(desc(2))
fa.sizes$V1 <- gsub("##contig=<ID=", "", fa.sizes$V1)
fa.sizes$V2 <- as.numeric(gsub(">", "", fa.sizes$V2))
colnames(fa.sizes) <- c("contig", "size")
# limit the data to only the 30 largest contigs
fa.sizes <- fa.sizes[1:min(nrow(fa.sizes), 30), ]
```

## General Information {data-height=100}
### nvariants
```{r echo=FALSE, message=FALSE, warning=FALSE}
summinfo <- sv %>%
  group_by(contig, type) %>%
  summarise(count = n()) %>% 
  group_by(type) %>%
  summarise(
    total = sum(count),
    avg_per_contig = round(mean(count), 2),
    sd = round(sd(count), 2)
)

nvar = sum(summinfo$total)
valueBox(scales::comma(nvar), caption = "Variants", color = ifelse(nvar > 0, "success", "warning"))
```

### nbreakends
```{r}
nbnd <- ifelse("BND" %in% summinfo$type, summinfo$type[which(summinfo$type == "BND")], 0)
valueBox(scales::comma(nbnd), caption = "Breakends", color = "info")
```
### ndeletions
```{r}
ndel <- ifelse("DEL" %in% summinfo$type, summinfo$type[which(summinfo$type == "DEL")], 0)
valueBox(scales::comma(ndel), caption = "Deletions", color = "info")
```

### nduplications
```{r}
ndup <- ifelse("DUP" %in% summinfo$type, summinfo$type[which(summinfo$type == "DUP")], 0)
valueBox(scales::comma(ndup), caption = "Duplications", color = "info")
```

### ninversions
```{r}
ninv <- ifelse("INV" %in% summinfo$type, summinfo$type[which(summinfo$type == "INV")], 0)
valueBox(scales::comma(ninv), caption = "Inversions", color = "info")
```

## Various Stats Headers
### summary info {.no-title}
<h2> Variant Information </h2>
Below are two tables with information on the variants. The first (left) table
details all the variants LEVIATHAN detected and the quantity/quality of evidence
for those detections. The second (right) table details each type of structural variant for
every contig. 

## Various Stats Tables
### All variants found
```{r echo = FALSE, warnings = FALSE, message = FALSE}
DT::datatable(
  sv,
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(dom = 'Brtip', buttons = c('csv'), scrollX = TRUE),
  autoHideNavigation = T,
  fillContainer = T
)
```

### Per-contig info
```{r echo = FALSE, message = FALSE}
sv$length <- gsub("\\.", "1", sv$length) %>%  as.numeric()
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
grpstats <- sv %>%
  group_by(contig, type) %>%
  summarise(count = n(), total_bp  = sum(length))

DT::datatable(
  grpstats, 
  rownames = F, 
  filter = "top", 
  extensions = 'Buttons', 
  options = list(dom = 'Brtip', buttons = c('csv'), scrollX = TRUE),
  autoHideNavigation = T,
  fillContainer = T
)
```

## SV size and location
```{r colors, echo = FALSE, warning = FALSE, message = FALSE}
col.palette <- c(
  "DEL" = "#5a8c84",
  "DUP" = "#ffd75f",
  "INV" = "#4a9fea",
  "BND" = "#df487f"
  )
```
### legend {.no-title data-width=200}
To the right is a circular plot to visualize the distribution of structural variants
across (up to) 30 of the largest contigs. This should help you assess
the presence/absence of expected variants. The legend below shows which
colors correspond to which type of variant.
You may click on the plot to expand it in your browser window. Clicking it again will exit
out of the zoomed view.

```{r}
justcolors <- c("#df487f", "#5a8c84", "#ffd75f", "#4a9fea")

data.frame(x = 1, y = 1:4, colour = c("Breakend","Deletion","Duplication", "Inversion")) %>% 
ggplot(aes(x, y, fill = colour))+
  geom_point(alpha=0, shape = 22, color = "white")+ # completely transparent rectangular point 
  scale_fill_manual(values=justcolors, drop=FALSE) +
  guides(fill = guide_legend(override.aes = list(alpha=1, size = 25)))+ # showing the point in the legend
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = c(0.3, 0.5),
        legend.text = element_text(size = 25),
        legend.title=element_text(size=25),
        legend.key = element_rect(fill='NA'),
        panel.grid = element_blank(),
        panel.border = element_rect(colour = "white", fill='white', linewidth=1)
  ) +
  labs(fill="Variant Type")
```

### Summary {.no-title data-height=700}
```{r echo = FALSE, message = FALSE, warning = FALSE, dev = "jpeg", out.width= "100%"}
circos.clear()
col_text <- "grey40"
circos.par("track.height" = 0.8, gap.degree = 3, cell.padding = c(0, 0, 0, 0))
circos.initialize(
  factors = fa.sizes$contig,
  xlim = matrix(c(rep(0, nrow(fa.sizes)), fa.sizes$size), ncol = 2)
)

# contigs
circos.track(
  ylim = c(0, 1),
  bg.col = "grey90",
  bg.border = F,
  track.height = 0.06,
  panel.fun = function(x, y) {
    chr = CELL_META$sector.index
    xlim = CELL_META$xlim
    ylim = CELL_META$ylim
    circos.text(
      mean(xlim),
      mean(ylim),
      chr,
      cex = 0.35,
      col = col_text,
      facing="bending.inside",
      niceFacing = T
    )
  }
)

# x axis
brk <- seq(0, 10, 0.5) * 10^7
circos.track(
  ylim = c(0,1),
  track.index = get.current.track.index(),
  bg.border = F,
  panel.fun = function(x, y) {
    circos.axis(
      h = "top",
      major.at = brk,
      labels = round(brk/10^6, 1),
      labels.cex = 0.4,
      col = col_text,
      labels.col = col_text,
      lwd = 0.7,
      labels.facing = "clockwise"
    )
  }
)

# INV
circos.genomicTrack(
  ylim = c(0,1),
  data = sv[sv$type == "INV", -1],
  track.height = 0.2,
  bg.border = F,
  panel.fun = function(region, value, ...) {
    circos.genomicRect(region, value, ybottom = 0, ytop = 1, col = "#4a9fea", border = "#4a9fea")
  }
)


circos.genomicTrack(
  ylim = c(0,1),
  data = sv[sv$type == "DEL", -1],
  track.height = 0.2,
  bg.border = F,
  panel.fun = function(region, value, ...) {
    circos.genomicRect(region, value, ybottom = 0, ytop = 1, col = "#5a8c84", border = "#5a8c84")
  }
)

circos.genomicTrack(
  ylim = c(0,1),
  data = sv[sv$type == "DUP", -1],
  track.height = 0.2,
  bg.border = F,
  panel.fun = function(region, value, ...) {
    circos.genomicRect(region, value, ybottom = 0, ytop = 1, col = "#ffd75f", border =  "#ffd75f")
  }
)

circos.genomicTrack(
  ylim = c(0,1),
  data = sv[sv$type == "BND", -1],
  track.height = 0.2,
  bg.border = F,
  panel.fun = function(region, value, ...) {
    circos.genomicRect(region, value, ybottom = 0, ytop = 1, col = "#df487f", border = "#df487f")
  }
)
```

# Per-Contig Plots
## contig plots
```{r}
color.palette <- c("DEL"="#5a8c84", "DUP"="#ffd75f", "INV"="#4a9fea","BND"="#df487f")
sv$ystart <- case_when(
  sv$type == "INV" ~ 0.1,
  sv$type == "DUP" ~ 1.1,
  sv$type == "DEL" ~ 2.1,
  sv$type == "BND" ~ 3.1,
)
sv$ystop <- sv$ystart + 0.8
```
### Per-contig {.no-title data-height=7000}
<h1> Structural Variants Per Contig </h1>
Below is a plot to help you assess what structural variants were detected by LEVIATHAN. These plots are interactive,
allowing you to hover over a variant to provide additional information, including the genomic interval in which it occurs and the
number of haplotag barcodes supporting the variant.

```{r}
l <- list()
nplotrows <- 0

for (i in 1:nrow(fa.sizes)) {
  sv.filt <- sv %>% filter(contig == fa.sizes$contig[i])
  sv_stats <- group_by(sv.filt, type) %>% summarise(n = length(type))
  if (nrow(sv_stats) == 0) {
    next
  }
  nplotrows <- nplotrows + 1
  plt <- sv.filt %>%
    ggplot() +
    geom_rect(
      alpha = 0.7,
      aes(
        ymin = ystart,
        ymax = ystop,
        xmin = position_start,
        xmax = position_end,
        fill = type,
        color = type,
        text = sprintf("Type: %s<br>Position: %s-%s<br>barcodes: %s", type, position_start, position_end, n_barcodes)
        )
      ) +
    geom_hline(yintercept = 1:3, color = "grey96") +
    scale_color_manual(values = color.palette) +
    scale_fill_manual(values = color.palette) +
    facet_grid(rows = vars(contig)) +
    theme_light() +
    theme(
      axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "grey80"),
      strip.text = element_text(color = "grey20"),
      legend.position = "none",
    ) +
    scale_x_continuous(labels = scales::comma) +
    coord_cartesian(xlim = c(0, fa.sizes$size[i] + 1), expand = F) +
    scale_y_continuous(
      breaks = c(0.5, 1.5, 2.5, 3.5),
      labels = c("INV", "DUP", "DEL", "BND"),
      limits = c(0,4)
    ) +
    xlab("Position (bp)")
    annotations <- list( 
      x = 0.5,
      y = 1.05,  
      text = as.character(fa.sizes$contig[i]),  
      xref = "paper",  
      yref = "paper",  
      xanchor = "center",  
      yanchor = "bottom",  
      showarrow = FALSE 
    )
    l[[i]] <- subplot(ggplotly(plt, height = 800, tooltip = "text"))
}
subplot(l, nrows = nplotrows, shareX = FALSE, titleX = TRUE)
```