---
title: "EMA Count Barcode Summary"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    vertical_layout: scroll
    favicon: "https://raw.githubusercontent.com/pdimens/harpy/docs/static/favicon.png"
---


```{r echo = FALSE, message = FALSE}
library(flexdashboard)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(magrittr)
library(DT)
library(plotly)
```

This report details the counts of valid and invalid haplotag barcodes in the sample read headers.
The correct haplotag format is `BX:Z:AxxCxxBxxDxx`, where there is a whitespace/tab before `BX:Z:` 
and `xx` is a 2-digit number between `00` and `96`. A valid FASTQ read header would then look like:
```
@A00814:267:HTMH3DRXX:2:1101:4580:1000 BX:Z:A02C01B11D46        RX:Z:GAAACGACCAACA+CGAACACGTTAGC   QX:Z:F,FFFFFFFFFFF+FF,FFF:FFFFFF
```
Notably, only the sequence ID (`@...`) and `BX:Z:` tag are required. In the example above, there are additional tags 
(`RX:Z:` and `QX:Z:`) which arent used by Harpy, but they conform to the 
[SAM comment spec (section 1.5)](https://samtools.github.io/hts-specs/SAMv1.pdf) of `TAG:TYPE:VALUE`. The takeaway 
is that the `BX:Z` tag can be anywhere in the read header after the sequence ID as long as any tags after it conform to the SAM spec `TAG:TYPE:VALUE`. 

```{r echo = FALSE, message = FALSE, warning = FALSE}
BXstats <- data.frame(
  Sample = character(),
  TotalBarcodes = integer(),
  ValidBarcodes = integer(),
  ValidPercent = numeric(),
  IgnoredReads = integer(),
  IgnoredPercent = numeric()
)

#for (i in list.files("~/Documents/harpy/test/Align/ema/logs/count/", ".count", full.names = TRUE)) {
for (i in snakemake@input[["countlog"]]) {
  samplename <- gsub(".count", "", basename(i))
  filetext <- readLines(i)
  bc_line <- gsub(",", "", filetext[2])
  matches <- gregexpr("[0-9]+", bc_line)
  counts <- as.numeric(unlist(regmatches(bc_line, matches)))
  bc_ok <- counts[1]
  bc_total <- counts[2]
  ignored <- as.numeric(gsub("[^0-9]", "", filetext[3]))
  if (bc_total == 0) {
    bc_percent <- 0
    ig_percent <- 0
  } else {
    bc_percent <- trunc((bc_ok / bc_total) * 10000) / 100
    ig_percent <- trunc((ignored / bc_total) * 10000) / 100
  }

  BXstats <- rbind(
    BXstats,
    data.frame(
      Sample = samplename,
      TotalBarcodes = bc_total,
      ValidBarcodes = bc_ok,
      ValidPercent = bc_percent,
      IgnoredReads = ignored,
      IgnoredPercent = ig_percent)
  )
}
```

## genstats {data-height=100}
### filerow {.no-title}
```{r}
valueBox(scales::comma(nrow(BXstats)), caption = "Samples", color = "success")
```
### mininv {.no-title}
```{r}
valueBox(min(BXstats$ValidPercent), caption = "Min Percent Valid", color = "info")
```
### maxinv {.no-title}
```{r}
valueBox(max(BXstats$ValidPercent), caption = "Max Percent Valid", color = "info")
```

### minign {.no-title}
```{r}
valueBox(min(BXstats$IgnoredPercent), caption = "Min Percent Ignored", color = "info")
```
### maxign {.no-title}
```{r}
valueBox(max(BXstats$IgnoredPercent), caption = "Max Percent Ignored", color = "info")
```

## Viz desc
### plot desc {.no-title}
Below is a plot summarizing the number of valid vs invalid barcodes per sample as
determined by `ema count` with a corresponding table
of those data with additional metrics. The plot is interactive and
hovering over the bars will display tooltips with the sample name and counts.

**Troubleshooting**: If for some reason `TotalBarcodes` equals `0`, then there may be an issue
with the format of your FASTQ headers. Please refer to the top of this report for
guidance on proper FASTQ read headers to hopefully remedy that issue.

## Per Sample Plot
### the plot {.no-title}
```{r echo = FALSE, message = FALSE, warning = FALSE}
plotdata <- BXstats %>%
  mutate(Invalid = TotalBarcodes - ValidBarcodes) %>%
  rename(Valid = ValidBarcodes) %>%
  select(Sample, Valid, Invalid)

percdata <- BXstats %>%
  mutate(Valid = round(ValidBarcodes / TotalBarcodes * 100, 2)) %>%
  mutate(Invalid = 100.0 - Valid) %>%
  select(Sample, Valid, Invalid)
```

```{r echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%'}
p <-  plotdata %>%
  pivot_longer(-Sample, names_to = "Barcode", values_to = "Count") %>%
  ggplot(aes(y = Sample, x = Count, fill = Barcode)) +
    geom_col() +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(), axis.title.y = element_blank()) +
    scale_fill_manual(values = c(Valid = "#afd5e7", Invalid = "#c02554")) +
    scale_x_continuous(labels = scales::comma) +
    labs(title = "Valid vs Invalid Barcodes")

pltheight <- 150 + (30 * length(unique(plotdata$Sample)))

ggplotly(p, height = pltheight)
```

### the table {.no-title}
```{r echo = FALSE, message = FALSE, warning = FALSE}
DT::datatable(
  BXstats,
  rownames = FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(dom = "Brtip", buttons = c("csv"), scrollX = TRUE),
  autoHideNavigation = T,
  fillContainer = T
)
```
