---
title: "FASTQ File Format Check"
date: "`r format(Sys.time(), '%m-%d-%y %X')`"
output:
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows  
    vertical_layout: scroll
    favicon: "https://raw.githubusercontent.com/pdimens/harpy/docs/static/favicon.png"
---


```{css zoom-lib-src, echo = FALSE, message = FALSE, warning = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.7.1/jquery.min.js"
```
```{js zoom-jquery, echo = FALSE, message = FALSE, warning = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '90%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r load environment, echo=FALSE, message=FALSE, warning=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
#infile <- "~/Documents/harpy/test/fastq/Preflight/filecheck.fastq.tsv"
infile <- snakemake@input[[1]]
dirname <- snakemake@params[[1]]
data <- read.table(infile, header = T)
attention_df <- data %>% 
  mutate(allmissing = (reads == noBX)) %>% 
  select(4:6) %>% 
  rowwise() %>% 
  summarise(count = sum(c(badBX,badSamSpec, allmissing)))
attention <- sum(attention_df$count > 0)
```
## filetop
### fltop {.no-title}
<h1> Directory: `r dirname` </h1>

This report reflects the FASTQ files identified in `r dirname`. Harpy has processed
the files in this folder to identify formatting issues that may require your attention.

## General Information {data-height=100}
### nfiles
```{r}
valueBox(scales::comma(nrow(data)), caption = "Files", color = "info")
```
### fileformat
```{r}
valueBox("FASTQ", caption = "File Format", color = "info")
```
### issues
```{r}
valueBox(scales::comma(attention), caption = "Files With Issues", color = ifelse(attention > 0, "warning", "success"))
```

## interpret
### interpcol {.no-title}
<h2> Interpreting the Data </h2>

The `harpy preflight fastq -d ...` created a `Preflight` folder within
the supplied directory and within it you will find this report, along with the
`filecheck.fastq.tsv` file summing the results that are included in this report.
This file contains a tab-delimited table with the columns: `file`,	`reads`, `noBX`,	`badBX`, and	`badSamSpec`. These columns are defined by:

- `file`: the name of the fastq file
- `reads`: the total number of reads in the file
- `noBX`: the number of reads that do not have a `BX:Z:` tag in the read header
  - severity: <strong style='color:#c0c0c0;'>minimal</strong> and likely won't cause issues unless no reads have `BX:Z:` tags
  - if you expect all of your reads should have `BX:Z:` tags, then further investigation is necessary
- `badBX`: the haplotag barcode in the `BX:Z:` comment does not adhere to the proper `AxxCxxBxxDxx` format
  - severity: <strong style='color:red;'>critical</strong> and will cause downstream issues and errors
- `badSamSpec`: the comments in the read header after the read ID do not conform to the `TAG:TYPE:VALUE` [SAM specification](https://samtools.github.io/hts-specs/SAMv1.pdf)
  - severity: <strong style='color:red;'>critical</strong> and will cause downstream issues and errors

### Proper header {.no-title}
<h2> Proper Read Headers </h2>

An example of a proper FASTQ read header is like the one below, where all comments
following the initial read ID (`@A00814...`) have the proper [SAM spec](https://samtools.github.io/hts-specs/SAMv1.pdf)
`TAG:TYPE:VALUE` format and among them is a `BX:Z:` tag followed by a `AxxCxxBxxDxx`
formatted haplotag barcode. The read header below contains the read ID `@A00814...`,
the `BX:Z:` haplotag barcode tag, and two more comments `RX:Z:` and `QX:Z:` that both
adhere to the SAM specification.

```
@A00814:267:HTMH3DRXX:2:1101:4580:1000 BX:Z:A02C01B11D46 RX:Z:GAAACGACCAACA+CGAACACGTTAGC QX:Z:F,FFFFFFFFFFF+FF,FFF:FFFFFF
```

## viz desc
### desc {.no-title}
<h2> FASTQ file format Check </h2>
Below are a plot and corresponding table of the results of the preflight checks. You
may click on the plot to expand it in your browser window. Clicking it again will exit
out of the zoomed view.

## datarow
### plot {.no-title}
```{r}
n <- nrow(data)
pltheight <- max(round(0.38 * n, digits = 0), 3)
```

```{r plottable, dev="jpeg", echo=FALSE, message=FALSE, warning=FALSE, fig.height=pltheight}
pivot_longer(data, -1, names_to = "Parameter", values_to = "Count") %>% 
ggplot(aes(x = Count, y = file, fill = Parameter, color = Parameter)) +
  geom_col(stat= "identity", position = position_dodge(width = 0.5), linewidth = 0.3) +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank()) +
  scale_fill_manual(values = c("indianred", "dodgerblue","orange","grey70")) +
  scale_color_manual(values = c("#a23c3c", "#2a7ed0","#dd8f00","grey60")) +
  #scale_color_manual(values = c("#d57575", "#369cff","#ffb42a","grey75")) +
  scale_x_continuous(labels = scales::comma) +
  ylab("") +
  labs(title = "FASTQ file check")
```

### table {.no-title}
```{r dtable, echo=FALSE, message=FALSE, warning=FALSE}
DT::datatable(data, rownames = F, filter = "top", extensions = 'Buttons', options = list(dom = 'Brtip', buttons = c('csv'), scrollX = T))
```
