---
title: "Imputation Assessment"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    vertical_layout: scroll
    favicon: "https://raw.githubusercontent.com/pdimens/harpy/docs/static/favicon_dark.png"
---

# General Stats
## header row
```{css zoom-lib-src, echo = FALSE, message = FALSE, warning = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.7.1/jquery.min.js"
```
```{js zoom-jquery, echo = FALSE, message = FALSE, warning = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '90%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r load environment, echo=FALSE, message=FALSE, warning=FALSE}
library(flexdashboard)
library(ggplot2)
library(magrittr)
library(tidyr)
library(dplyr)
library(DT)
library(scales)
modelparams <- paste(snakemake@params[[1]])
#modelparams <- "Filler Text"
```
```{r echo = FALSE, message = FALSE, warning = FALSE}
comparefile <- snakemake@input[[1]]
#comparefile <- "~/impute.compare.stats"
tryCatch(
  expr = {
    gcts <- read.table(comparefile, header = F)[, c(2, 23, 25, 24, 27)]
  },
  error = function(e){ 
    print("Stats file is empty. Perhaps there were no SNPs that bcftools identified.")
    knitr::knit_exit()
  }
)
names(gcts) <- c("sample","RefHom", "Het",	"AltHom", "missing")

conv_height <- 1.4 + (0.12 * nrow(gcts))
#frameheight <- 1.2 + (0.166 * nrow(invalids))

gcts_long <- pivot_longer(gcts, -1, names_to = "Conversion", values_to = "Count")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
infofile <- snakemake@input[[2]]
#infofile <- "~/impute.infoscore"

tryCatch(
  expr = {
    infos <- read.table(infofile, header = F, col.names = c("contig", "position", "info"))
  },
  error = function(e){ 
    print("Info score file is empty. Perhaps there were no SNPs that bcftools identified.")
    knitr::knit_exit()
  }
)
```

### reportheader {.no-title}
<h2> `r comparefile` </h2>
This report details the results of genotype imputation using STITCH with the model parameters:
`r modelparams`.

This page displays various information and general statistics and clicking the `Individual Statistics` tab 
in the navigation menu above will display imputation results per individual.

## General Information {data-height=100}
```{r}
gcts_minmax <- filter(gcts_long, Conversion != "missing") %>% summarise(min = min(Count), max = max(Count))
```

### nsamples
```{r}
valueBox(scales::comma(nrow(gcts)), caption = "Samples", color = "success")
```

### min impute
```{r}
valueBox(scales::comma(gcts_minmax$min), caption = "Minimum Imputations", color = "info")
```
### maximpute
```{r}
valueBox(scales::comma(gcts_minmax$max), caption = "Maximum Imputations", color = "info")
```

### mean infoscore
```{r}
valueBox(round(mean(infos$info), digits = 2), caption = "Average INFO_SCORE", color = "info")
```

### sd infoscore
```{r}
valueBox(round(sd(infos$info), digits = 2), caption = "StDev INFO_SCORE", color = "info")
```


## INFO Scores
### INFO desc {.no-title}
<h3> Distribution of INFO_SCORE </h3>
`STITCH` outputs a novel `INFO/INFO_SCORE` tag in the resulting VCF files that represents the
"quality" of the imputation. This value ranges from `0` to `1.0`, with `0` being
very poor, and `1.0` being very good. You will want to filter this file to remove
genotypes with a low `INFO_SCORE`. Below is the frequency distribution of the global
and per-contig `INFO_SCORE` values for the variants in the VCF file.  The  per-contig
plot shows up to 30 of the largest contigs present in the VCF file. Clicking
on a plot will expand it to fill your browser window. Clicking it again will exit
out of the zoomed view.

## distributions
### Overall distribution {.no-title}
```{r echo=FALSE, message=FALSE, warning=FALSE}
maxsizes <- infos %>% group_by(contig) %>% summarise(maxsize = max(position)) %>% arrange(desc(maxsize))
maxsizes <- maxsizes[1:min(nrow(maxsizes), 30), ]

ncontig <- nrow(maxsizes)
pltheight <- round(1.1 * ncontig, digits = 0)
```

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=2.5, out.width= "100%"}
ggplot(infos, aes(x = info)) +
  geom_histogram(fill = "grey60", color = "grey40", linewidth = 0.25, bins = 100, boundary = 0.5) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank()
  ) +
  scale_y_continuous(labels = scales::comma) +
  xlab("INFO_SCORE") +
  ylab("Count") +
  labs(title = "Distribution of imputation INFO_SCORE")
```

### Per-contig distribution {.no-title}
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=pltheight, fig.align='center', out.width = "100%"}
ggplot(infos[infos$contig %in% maxsizes$contig, ], aes(x = info)) +
  geom_histogram(fill = "grey60", color = "grey40", linewidth = 0.25, bins = 100, boundary = 0.5) +
  facet_wrap(~contig, ncol = 2, scales = "free_y") +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank()
  ) +
  scale_y_continuous(labels = comma) +
  xlab("INFO_SCORE") +
  ylab("Count") +
  labs(title = "Distribution of imputation INFO_SCORE")
```

## filtering
### Filtering Expectations desc {.no-title}
<h3> Filtering Expectations </h3>
You will want to filter the resulting variants
based on this `INFO/INFO_SCORE` tag, which can be done that simply with `bcftools`.
Here is an example snippet of keeping loci with an `INFO_SCORE` of at least `0.2`: 
```
# use -Ob to output to bcf instead of vcf to save on disk space

bcftools view -i 'INFO/INFO_SCORE >= 0.2' -Ob file.vcf > filtered.file.bcf
```

Below is a table of the number of variants sites that may remain if filtering
at different minimum `INFO/INFO_SCORE` thresholds. Each column after `Total Variants`
refers to a filtering threshold and the numbers within that column are the number of 
SNPs that would be kept if filtering out SNPs with an `INFO/INFO_SCORE` value below 
it (_i.e._ `0.2` is the number of SNPs with a score `>= 0.2` ).

## filtering table
### filt table {.no-title}
```{r echo = FALSE, message = FALSE, warning = FALSE}
filterthresh <- infos %>% group_by(contig) %>%
  summarise(
    totals = length(info),
    twenty = sum(info >= 0.2),
    twentyfive = sum(info >= 0.25),
    thirty = sum(info >= 0.30),
    forty = sum(info >= 0.40),
    fifty = sum(info >= 0.50),
    seventyfive = sum(info >= 0.75),
    ninety = sum(info >= 0.90),
    )

tots <- colSums(filterthresh[, -1])
filterthresh <- rbind(c("all", tots), filterthresh) 
if nrow(filterthresh) < 1 {
  cat("No variants retained")
} else {
  DT::datatable(
    filterthresh,
    rownames = FALSE,
    extensions = "Buttons",
    colnames = c("Contigs", "Total Variants", "0.20", "0.25", "0.30", "0.4", "0.5", "0.75", "0.90"),
    options = list(
      dom = "Brtip",
      buttons = c("csv"),
      scrollX = TRUE
    ),
    autoHideNavigation = T,
    fillContainer = T
  )
}
```

# Individual Statistics
## Ind stats desc
### plotdesc {.no-title}
<h3> Imputations Per Sample </h3>
This chart visualizes the count and proportion of `missing` genotypes per sample imputed
into other genotypes. Following standard VCF convention, `Ref` refers to
the reference allele (from the reference genome), `Alt` refers to the alternative 
allele, and `Missing` details how many SNPs remain missing
(failed to impute) for that individual. As an example, the `Homozygous Ref` is 
the number of missing genotypes that were imputed into
a homozygous genotype for the reference allele. The table on the right is the
underlying data used to create the visualization. Clicking
on a plot will expand it to fill your browser window. Clicking it again will exit
out of the zoomed view.

# Individual Statistics
## ind stats
### Individuals Plot {.no-title}
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height=conv_height, out.width= "100%"}
ggplot(gcts_long, aes(y = sample, x = Count, fill = Conversion)) +
  geom_col(position = "stack") +
  scale_x_continuous(expand = c(0,0), labels = comma) +
  ylab("") +
  theme_minimal() +
  theme(panel.grid.major.y = element_blank(), axis.ticks.y = element_line(size = 0.5)) +
  scale_fill_manual(
    labels = c("→ Homozygous Alt", "→ Heterozygote", "→ Missing", "→ Homozygous Ref"),
    values = c("#93d2fa", "#83afa8", "#424043", "#ffd75f")
  ) +
  labs(title = "Missing Genotype Imputations")
```

### Imputation Results Table {.no-title}
```{r success table, echo=FALSE, message=FALSE, warning=FALSE}
DT::datatable(
  gcts,
  rownames = FALSE,
  extensions = "Buttons",
  colnames = c("Sample", "Homozygous Ref", "Homozygous Alt", "Heterozygote", "Missing"),
  options = list(
    dom = "Brtip",
    buttons = c("csv"),
    scrollX = TRUE
  ),
  autoHideNavigation = T,
  fillContainer = T
)
```