---
title: "BAM Haplotag Format Check"
date: "`r format(Sys.time(), '%d %b, %Y at %H:%M')`"
output:
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows  
    vertical_layout: scroll
    logo: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_report.png
    favicon: "https://raw.githubusercontent.com/pdimens/harpy/docs/static/favicon_report.png"
    navbar:
      - { title : "Docs", icon: "fa-book", href: "https://pdimens.github.io/harpy/", align: right }
---

```{r load environment, echo=FALSE, message=FALSE, warning=FALSE}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("flexdashboard","dplyr","tidyr","DT","highcharter","scales")

#infile <- "~/bamfilecheck.bam.tsv"
infile <- snakemake@input[[1]]
data <- read.table(infile, header = T)
attention_df <- data %>% select(-1,-3) %>% rowwise() %>% summarise(count = sum(c(nameMismatch, noBX, badBX)))
attention <- sum(attention_df$count > 0)
```
# File Checks
## filetop
### fltop {.no-title}
<h1> Preflight checks for BAM files </h1>

This report reflects the BAM files provided. Harpy has processed
the files to identify formatting conformity and issues that may require your attention.

## General Information {data-height=100}
### nfiles
```{r}
valueBox(scales::comma(nrow(data)), caption = "Alignment Files", icon = "fa-file", color = "#C0C0C0")
```

### no MI tag
```{r}
noMItag <- sum(data$noMI ==data$alignments)
valueBox(scales::comma(noMItag), caption = "No MI: tag", icon = ifelse(noMItag > 0, "fa-exclamation", "fa-check"), color = ifelse(noMItag > 0, "warning", "success"))
```

### bxnotlast
```{r}
bxnotlast <- sum(data$bxNotLast > 0)
valueBox(scales::comma(bxnotlast), caption = "BX:Z: tag not last", icon = ifelse(bxnotlast > 0, "fa-exclamation", "fa-check"), color = ifelse(bxnotlast > 0, "warning", "success"))
```

### issues
```{r}
valueBox(scales::comma(attention), caption = "Files With Issues", icon = ifelse(attention > 0, "fa-exclamation", "fa-check"), color = ifelse(attention > 0, "warning", "success"))
```

## viz desc
### desc {.no-title}
<h2> BAM file format Check </h2>
Below are a plot and corresponding table of the results of the preflight checks.
The plot renames the parameters for clarity and they are:

```{r}
knitr::kable(
  data.frame(
    "data column" = c("`nameMismatch`", "`noMI`", "`noBX`", "`badBX`", "`bxNotLast`"),
    "name in plot" = c("name matches", "MI: tag", "BX:Z: tag", "AxxCxxBxxDxx format", "BX:Z: last tag"),
    "pass condition" = c(
      "the file name matches the `@RG ID:` tag in the header",
      "**any** alignments with `BX:Z:` tags also have `MI:i:` (or `MI:Z:`) tags",
      "any `BX:Z:` tags present",
      "**all** alignments with BX:Z: tag have properly formatted `AxxCxxBxxDxx` barcodes",
      "**all** reads have `BX:Z`: as final tag in alignment records"
      ),
    "fail condition" = c(
        "file name does not match `@RG ID:` in the header",
        "**all** reads have `BX:Z:` tag present but `MI:i:` tag absent",
        "**all** alignments lack `BX:Z:` tag",
        "**any** `BX:Z:` barcodes have incorrect format",
        "**at least 1 read** doesn't have `BX:Z:` tag as final tag"
      )
  ), col.names = c("data column", "name in plot", "pass condition", "fail condition")
)
```

## datarow
### plot {.no-title}
```{r status_df}
data2 <- data %>% 
  mutate(
    nameMismatch = ifelse(nameMismatch > 0, F, T),
    noMI = ifelse(noMI == alignments, F, T),
    noBX = ifelse(noBX == alignments, F, T),
    badBX = ifelse(badBX > 0, F, T),
    bxNotLast = ifelse(bxNotLast > 0, F, T)
    ) %>% 
  select(-3)
m <- as.matrix(data2[,-1]) * 1
colnames(m) <- c("Name matches", "MI: tag present", "BX:Z: tag present", "AxxCxxBxxDxx format", "BX:Z: last tag")
rownames(m) <- gsub(".bam", "", data$file)
```

```{r readsper, echo = FALSE, message = FALSE, warning = FALSE, out.width = '100%'}
hchart(m) |>
  hc_colorAxis(stops = list(c(0, "#feda75"), c(1, "#4faad1")), tickAmount = 2, min = 0, max = 1) |>
  hc_title(text = "Alignment File Checks") |>
  hc_xAxis(labels = list(style = list(fontSize = '16px')) )|>
  hc_yAxis(labels = list(style = list(fontSize = '16px')) )|>
  hc_plotOptions(heatmap = list(animation = FALSE, borderWidth = 1, borderColor = "white", states = list(inactive = list(animation = FALSE), normal = list(animation = FALSE)))) |>
  hc_legend(enabled = F) |>
  hc_tooltip(
    animation = FALSE,
    formatter = JS("function () {return '<b>' + this.series.xAxis.categories[this.point.x] + '</b><br>' + this.series.yAxis.categories[this.point.y] + '<br/>QC: <b>' + (this.point.value<1 ? 'FAIL</b>' : 'PASS</b>');}") 
  ) |>
  hc_exporting(enabled = T, filename = "bam.preflight",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

### table {.no-title}
```{r dtable, echo=FALSE, message=FALSE, warning=FALSE}
DT::datatable(
  data,
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(dom = 'Brtip', buttons = c('csv'), scrollX = T),
  fillContainer = T
)
```


# Details
## interpret
### interpcol {.no-title}
<h2> Interpreting the Data </h2>

The `harpy preflight bam ...` command created a `filecheck.bam.tsv` file in the specified
output directory that summarizes the results that are included in this report.
This file contains a tab-delimited table with the columns: `file`,	`nameMismatch`, `alignments`,	`noBX`, and	`badBX`. These columns are defined by:

#### file
The name of the BAM file

#### nameMismatch
The sample name of the file inferred from name of the file (i.e. Harpy assumes `sample1.bam` to be `sample1`) does not match the `@RG ID` tag in the alignment file.

- severity: <strong style='color:red;'>critical</strong> and will likely cause downstream issues and errors

#### alignments
The total number of alignment records in the file

#### noMI
Alignment records lack `MI:` tag (`MI:Z:` or `MI:i:`)

- severity: <strong style='color:red;'>critical</strong> and will prevent HapCut2 phasing
- the `MI` tag is the "molecular identifier", meaning it's the unique name of a molecule from which are series of alignments share a barcode. 
  - `EMA` adds these to alignments, `BWA` does not, but `harpy align` will add them when using `BWA`

#### noBX
The number of alignments that do not have a `BX:Z:` tag in the record

- severity: <strong style='color:#c0c0c0;'>minimal</strong> and likely won't cause issues unless no alignments have `BX:Z:` tags
- if you expect all of your alignments should have `BX:Z:` tags, then further investigation is necessary

#### badBX
The haplotag barcode in the `BX:Z:` tag does not adhere to the proper `AxxCxxBxxDxx` format

- severity: <strong style='color:red;'>critical</strong> and will cause downstream issues and errors

#### bxNotLast
The `BX:Z:` tag in the alignment record is not the last tag

- severity: <strong style='color:orange;'>conditional</strong> and only relevant for LEVIATHAN variant calling 
- can be ignored if not intending to call structural variants with LEVIATHAN
- if intending to use LEVIATHAN, the `BX:Z:` tag must be the last tag in the alignment records
  - can be fixed by editing the BAM file, or fixing FASTQ files prior to alignment
  - `harpy align -m bwa` automatically moves the `BX:Z:` tag to the end of the alignment record


### Proper format {.no-title}
<h2> Proper BAM file format </h2>

#### BX Tag
A proper linked-read alignment file will contain a `BX:Z:` tag with an alignment record
that features a properly-formatted haplotag barcode `AxxCxxBxxDxx`. If this barcode is not in that
format, then it's likely the input FASTQ used for read mapping is the source of the
issue. You can check those FASTQ files for errors with `harpy preflight fastq ...`. 

#### Example Alignment Record
Below is an example of a proper alignment record for a file named `sample1.bam`.
Note the tag `RG:Z:sample1`, indicating this alignment is associated with `sample1` and
matches the file name. Also note the correctly formatted haplotag barcode `BX:Z:A19C01B86D78`
and the presence of a `MI:` tag. To reduce horizontal scrolling, the alignment sequence and Phred
scores have been replaced with `SEQUENCE` and `QUALITY`, respectively.

```
A00814:267:HTMH3DRXX:2:1132:26268:10316 113     contig1 6312    60      4S47M1D86M      =       6312    0       SEQUENCE  QUALITY     NM:i:2  MD:Z:23C23^C86  MC:Z:4S47M1D86M AS:i:121        XS:i:86 RG:Z:sample1   MI:i:4040669       RX:Z:GAAACGATGTTGC+CCTAAGCCGAATC        QX:Z:FFFFFFFFFFFFF+FFFFF:FFFFFFF    BX:Z:A19C01B86D78
```