---
title: "Harpy Alignment Summary"
date: "`r format(Sys.time(), '%d %b, %Y at %H:%M')`"
output:
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    vertical_layout: scroll
    horizontal_layout: fill
    mathjax: NULL
    logo: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_report.png
    favicon: "https://raw.githubusercontent.com/pdimens/harpy/docs/static/favicon_report.png"
    navbar:
      - { title : "Docs", icon: "fa-book", href: "https://pdimens.github.io/harpy/", align: right }
      - { title : "Source", icon: "fa-github", href: "https://www.github.com/pdimens/harpy/", align: right }
---

# Barcode Stats
```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
logfile <- file(snakemake@log$logfile, open = "wt")
sink(logfile)
sink(logfile, type = "message")
```

```{r package_imports, results = F}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}

using("flexdashboard","dplyr","highcharter","DT","BioCircos")
```

```{r import_file, results = F}
#infile <- "~/file.bxstats.gz"
# samplename <- "test sample"
infile <- snakemake@input$bxstats
samplename <- snakemake@params$samplename
bamfile <- paste0(samplename, ".bam")
tb <- read.table(infile, header = T, sep = "\t") %>% select(-start, -end)
if(nrow(tb) == 0){
  print(paste("Input data file",infle, "is empty"))
  knittr::knit_exit()
}
tb$valid <- tb$molecule != "invalidBX"
```

```{r bxper, results = F}
valids <- filter(tb, valid)
invalids <- filter(tb, !valid)
valids$molecule <- as.integer(valids$molecule)
nBX <- group_by(valids, contig) %>% 
  summarize(nBX = length(molecule))

avgBX <- round(mean(nBX$nBX), digits = 2)
totuniqBX <- read.table(infile, header = F, sep = "\n", as.is = T, skip = nrow(tb) + 1, comment.char = "+")
totuniqBX <- gsub("#total unique barcodes: ", "", totuniqBX) |> as.integer()

tot_invalid <- sum(tb[!(tb$valid), "reads"])
tot_valid <- sum(tb[tb$valid, "reads"])
```

```{r nxx, results = F}
NX <- function(lengths, X=50){
  lengths <- as.numeric(sort(lengths, decreasing=TRUE))
  index <- which(cumsum(lengths) / sum(lengths) >= X/100)[1]
  return(lengths[index])
}
```

## fileheader
### hdr {.no-title}
<h1> Linked-Read Barcode Statistics </h1>
The information presented below were gathered from the alignments within **`r bamfile`**.


## General Information {data-height=100}
### ncontigs
```{r valuebox_ncontigs}
valueBox(scales::comma(length(unique(tb$contig))), caption = "Contigs", color = "#ffffff")
```

### glob-total-bx
```{r valuebox_n_barcodes}
valueBox(scales::comma(totuniqBX), caption = "Unique barcodes", color = "#8f4f76")
```

### mol-thresh
```{r valuebox_mol_thresh}
if(snakemake@params$mol_dist > 0){
  valueBox(scales::comma(snakemake@params$mol_dist), caption = "Molecule threshold", color = "#be96b9")
}
```

### glob-total-mol
```{r valuebox_n_molecules}
valueBox(scales::comma(length(unique(valids$molecule))), caption = "Unique Molecules")
```

### validBX
```{r valuebox_valids}
valueBox(scales::comma(tot_valid), caption = "Valid BX Records", color = "success", icon = "fa-vial-circle-check")
```

### invalidBX
```{r valuebox_invalids}
valueBox(scales::comma(tot_invalid), caption = "Invalid BX Records", color = "warning", icon = "fa-x")
```

### glob-avg
```{r valuebox_mol_per_contig}
valueBox(scales::comma(avgBX), caption = "Avg molecules per contig", color = "#d4d4d4")
```


## N50 and N90
### Molecule NXX Length Metrics
```{r NX_stats}
summary_table <- valids %>% 
    group_by(contig) %>%
    summarize(valid_records = sum(reads), molecules = length(molecule),n50 = NX(length_inferred, 50), n75 = NX(length_inferred, 75), n90 = NX(length_inferred, 90))
summ_invalid <- invalids %>% group_by(contig) %>% summarize(invalid_records = sum(reads))
summary_table <- left_join(summary_table, summ_invalid, by = "contig", keep = T) %>% select(1, 2, 8, 3:6)
summary_table$invalid_records[is.na(summary_table$invalid_records)] <- 0
DT::datatable(
  summary_table,
  rownames = F,
  options = list(
    dom = 'Brtip',
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_molecule_NX")))
  ),
  colnames = c("contig", "valid BX records", "invalid BX records", "molecules", "N50", "N75", "N90"),
  fillContainer = T
)
```

## Reads per molecule dec
### Reads per mol dec {.no-title}
<h2> Reads per Molecule </h2>
The chart below shows the distribution of the number of reads per haplotag molecule. That is, how many alignments
are associated with a unique molecule. This excludes the number of reads associated with invalid
or absent haplotag barcodes.

### bases per desc {.no-title}
<h2> Total Bases Aligned </h2>
Below is a frequency distribution showing the number of base pairs aligned
per unique molecule. These data are shown in 500 bp bins. Pay attention to the Y axis,
as it likely doesn't start at `0`.

## the plots
### reads per {.no-title}

```{r readsper, out.width = '100%'}
hs <- hist(
  valids$reads,
  breaks = min(valids$reads):max(valids$reads),
  plot = F
)
hs$counts <- round(hs$counts / sum(hs$counts) * 100, 4)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)

hchart(hs, "areaspline", hcaes(x = val, y = freq), color = "#8484bd", name = "% of molecules", marker = list(enabled = FALSE)) |>
  hc_title(text = "Reads Per Molecule") |>
  hc_xAxis(title = list(text = "reads per molecule")) |>
  hc_yAxis(title = list(text = "% molecules")) |>
  hc_caption(text = paste0("Total unique molecules: ", totuniqBX)) |>
  hc_tooltip(crosshairs = TRUE) |>
  hc_exporting(enabled = T, filename = paste0(samplename, ".readsper"),
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

### bases per {.no-title}
```{r basesper, out.width="100%"}
hs <- hist(round(valids$aligned_bp, -2), breaks = 50,  plot = F)
hs$counts <- round(hs$counts / sum(hs$counts)*100,4)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)

hchart(hs, "areaspline", hcaes(x = val, y = freq), color = "#75b89e", name = "% of molecules", marker = list(enabled = FALSE)) |>
  hc_title(text = "Bases Aligned Per Molecule") |>
  hc_xAxis(title = list(text = "aligned bases per molecule")) |>
  hc_yAxis(title = list(text = "% molecules")) |>
  hc_caption(text = paste0("Total unique molecules: ", totuniqBX)) |>
  hc_tooltip(crosshairs = TRUE) |>
  hc_exporting(enabled = T, filename = paste0(samplename, ".basesper"),
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

## inferred-header
### inferred desc {.no-title}
<h2> Inferred Molecule Lengths </h2>
The chart below shows the frequency distribution of molecule lengths 
inferred from the first and last alignment positions along a contig for all
alignments associated with a single haplotag barcode on a given contig. 
Pay attention to the Y axis, as it may not start at `0`.

## inferred-plot
### inferredplot {.no-title}
```{r inferred, out.width = '100%'}
hs <- hist(
  round(valids$length_inferred / 1000,0),
  breaks = 25,
  plot = F
)
hs$counts <- round(hs$counts / sum(hs$counts)*100,2)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)

hchart(hs, "areaspline", hcaes(x = val, y = freq), color = "#b3519d", name = "% of molecules", marker = list(enabled = FALSE)) |>
  hc_title(text = "Inferred Molecule Length") |>
  hc_subtitle(text = "lengths reported as kilobases (kbp)") |>
  hc_xAxis(title = list(text = "Inferred Molecule length (kbp)"), type = "logarithmic") |>
  hc_yAxis(title = list(text = "% of molecules")) |>
  hc_caption(text = paste0("Total unique molecules: ", totuniqBX)) |>
  hc_tooltip(crosshairs = TRUE) |>
  hc_exporting(enabled = T, filename = paste0(samplename, ".mollen"),
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

## breadth
### inferred_cov desc {.no-title}
<h2> Inferred Molecule Read Coverage </h2>
The charts below show the frequency distribution of molecule coverage, meaning
the percent of each inferred molecule that has sequences aligned to it. In other
words, "how much of each unique long molecule is actually sequenced/aligned?"
Keep in mind there is a disctinction between sequences and alignments, since
some sequences belonging to a particular molecule may not align well and wouldn't
appear in the alignment data. The chart on the left shows molecule coverage as computed
by the number of bases aligned to that molecule. The chart on the right shows molecule
coverage as computed by the inferred fragment length from which the aligned sequences originate.
In other words, this computation considers the length of the original DNA fragment that was put
on the sequencer, which may be longer than the 150bp (x2) inserts that actually got sequenced.

If it helps, imagine a 500bp haplotagged DNA fragment being sequenced as a 150bp paired-end
sequence and all 300bp aligned to your reference genome. The plot on the left calculates the proportion of
molecule coverage as `300bp รท molecule_length`, whereas the plot on the right calculates this proportion as
`500bp รท molecule_length`, which is the 300bp that aligned + the 200bp between the paired-end reads that were not sequenced.

## breadthplot
### coverage by aligned bp {.no-title}
```{r inferred_cov_bp, out.width = '100%'}
hs <- hist(
  round(valids$coverage_bp, 0),
  breaks = seq(0, 1, by = 0.05),
  plot = F
)
hs$counts <- round(hs$counts / sum(hs$counts)*100,4)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)

hchart(hs, "areaspline", hcaes(x = val, y = freq), color = "#e59765", name = "% of molecules", marker = list(enabled = FALSE)) |>
  hc_title(text = "Percent Molecule Coverage by Aligned Bases") |>
  hc_xAxis(title = list(text = "% molecule covered")) |>
  hc_yAxis(title = list(text = "% of molecules")) |>
  hc_caption(text = paste0("Total unique molecules: ", totuniqBX)) |>
  hc_tooltip(crosshairs = TRUE) |>
  hc_exporting(enabled = T, filename = paste0(samplename, ".molcov"),
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

### coverage by inserts {.no-title}
```{r inferred_cov_insert, out.width = '100%'}
hs <- hist(
  round(valids$coverage_inserts, 0),
  breaks = seq(0, 1, by = 0.05),
  plot = F
)
hs$counts <- round(hs$counts / sum(hs$counts)*100,2)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)

hchart(hs, "areaspline", hcaes(x = val, y = freq), color = "#dbe465", name = "% of molecules", marker = list(enabled = FALSE)) |>
  hc_title(text = "Percent Molecule Coverage by Inferred Insert Size") |>
  hc_xAxis(title = list(text = "% molecule covered")) |>
  hc_yAxis(title = list(text = "% of molecules")) |>
  hc_caption(text = paste0("Total unique molecules: ", totuniqBX)) |>
  hc_tooltip(crosshairs = TRUE) |>
  hc_exporting(enabled = T, filename = paste0(samplename, ".molcov"),
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

## Interpreting the supporting data {.data-height=50}
### interp desc {.no-title}
<h2> Interpreting the Data </h2>
These descriptions should help you understand the underlying data.

## inttable
### interpreting {.no-title}
<h3> Interpreting the Supporting File </h3>
Listed below are the descriptions of the columns in **`r basename(infile)`**, which was created
by Harpy using the included `bx_stats.py` script. The term `molecule` refers to the
`MI:i` tag in the alignments, which is a unique molecule ID given to the original
fragment alignments sharing a barcode are inferred to have originated from. The inference
takes into account an [alignment distance threshold](https://pdimens.github.io/harpy/haplotagdata/#barcode-thresholds)
and that the sequences aligned to the same contig.

```{r cols_explained}
knitr::kable(
  data.frame(
    "Column Name" = c("contig", "molecule", "reads", "start", "end", "length_inferred", "percent_coverage", "aligned_bp"),
    "Description" = c(
      "name of the contig the molecule occurs on",
      "the molecule name as given by the MI:i: tag",
      "number of alignments associated with this molecule",
      "the start position of the first alignment for that molecule",
      "the end position of the last alignment for that molecule",
      "inferred length of the molecule based on the start/end of the alignments sharing the same barcode",
      "what percent of the molecule is represented by sequence alignments",
      "total number of base pairs aligned for that molecule"
      )
    )
)
```

### Barcode validity {.no-title}
<h3> Interpreting Barcode Validity </h3>
BX barcode validity is classified into one of two categories:

```{r bx_explanation}
knitr::kable(
  data.frame(
    "Classification" = c("valid BX", "invalid BX"),
    "Description" = c(
      "a complete BX barcode was present in the read (i.e. no 00 for any segments)",
      "a barcode was present in the read, but it contained 00 in at least one of the barcode segments"
      )
    )
)
```

# Coverage Stats
```{r reset_env}
# clear out objects in memory to reduce RAM usage
rm(hs)
rm(valids)
rm(invalids)
```
```{r imports}
covfile <- snakemake@input$coverage
molcovfile <- snakemake@input$molecule_coverage
windowsize <- as.integer(snakemake@params$window_size)
samplename <- snakemake@params$samplename
#samplename <- "test sample"
#windowsize <- 50000
#covfile <- "~/cov.gz"
#molcovfile <- "~/molcov.gz"
tb <- read.table(covfile, header = F)
if(nrow(tb) == 0){
  print(paste("Input data file", covfile, "is empty"))
  knittr::knit_exit()
}
tbmol <- read.table(molcovfile, header = F)
if(nrow(tbmol) == 0){
  print(paste("Input data file", molcovfile, "is empty"))
  knittr::knit_exit()
}
```

```{r setup_cov_table}
colnames(tb) <- c("contig", "position", "depth")
colnames(tbmol) <- c("contig", "position", "mol_depth")
tb <- merge(tb, tbmol, by = c("contig", "position"))
q99 <- quantile(tb$depth, 0.99)
mol_q99 <- quantile(tb$mol_depth, 0.99)
windowskb <- round(windowsize/1000, 0)
rm(tbmol)
```

```{r summary_stats}
global_avg <- mean(tb$depth)
global_sd <- sd(tb$depth)
mol_global_avg <- mean(tb$mol_depth)
mol_global_sd <- sd(tb$mol_depth)

tb$outlier <- tb$depth > q99
outliers <- tb[tb$outlier, -5]
nonoutliers <- tb[!(tb$outlier), -5]

contig_avg <- group_by(tb, contig) %>%
  summarize(average = mean(depth), mol_average = mean(mol_depth), sdv = sd(depth), mol_sdv = sd(mol_depth))
contig_avg <- rbind(data.frame(contig = "global", average = global_avg, sdv = global_sd, mol_average = mol_global_avg, mol_sdv = mol_global_sd), contig_avg) %>% 
  mutate(average = round(average, 2), sdv = round(sdv, 2), mol_average = round(mol_average, 2), mol_sdv = round(mol_sdv, 2))
```


## covfileheader
### hdrcov {.no-title}
<h1> Alignment Coverage Statistics </h1>

This report contains information regarding the sequence alignment coverage
and depth for the file **`r paste0(samplename, ".bam")`**. The term `<Q99` here and
elsewhere in this report refers to keeping intervals whose depth is below  the 99th
depth percentile (`r q99`). The Q99 described is shown for diagnostic purposes, the
alignments above this depth were not removed from the file.

## General Information {data-height=100}
### ncontigs
```{r valuebox_contigs}
valueBox(scales::comma(length(unique(tb$contig))), caption = "Contigs", color = "success")
```

### general-windowsize
```{r valuebox_intervals}
valueBox(paste(windowskb, "kb"), caption = "Intervals", color = "info")
```
### glob-avg
```{r valuebox_glob_avg}
valueBox(round(global_avg,2), caption = "Average depth", color = "info")
```

### glob-sd
```{r valuebox_stdv}
valueBox(round(global_sd,2), caption = "Stdev depth", color = "info")
```

### mol-avg
```{r valuebox_mol_avg}
valueBox(round(mol_global_avg,2), caption = 'Mol. Average Depth', color = "info")
```

### mol-sd
```{r valuebox_mol_stdv}
valueBox(round(mol_global_sd,2), caption = 'Mol. Stdev depth', color = "info")
```

## Distdesc header
### distdesc {.no-title}
<h2> Alignment Depth Distribution </h2>
Below are the frequencies of interval coverage across all **`r windowskb` kilobase** intervals for all contigs.
For visual clarity, the X-axis of this plot is truncated at the 99% quantile of depth
values, which is **`r q99`** for these data.

## distributionplot
### distplot {.no-title}
```{r depth_distribution}
hs <- hist(tb$depth[tb$depth <= q99], breaks = 30, plot = F)
hs$counts <- round(hs$counts / sum(hs$counts) * 100, 2)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)
mol_hs <- hist(tb$mol_depth[tb$mol_depth <= mol_q99], breaks = 30, plot = F)
mol_hs$counts <- round(mol_hs$counts / sum(mol_hs$counts) * 100, 2)
mol_hs <- data.frame(val = mol_hs$breaks[-1], freq = mol_hs$counts)


highchart() |>
  hc_chart(type = "area", animation = F) |>
  hc_add_series(data = hs, type = "areaspline", color = "#56486d", name = "by alignments", hcaes(x = val, y = freq),  marker = list(enabled = FALSE)) |>
  hc_add_series(data = mol_hs, type = "areaspline", hcaes(x = val, y = freq), color = "#88c793", name = "by inferred molecules", marker = list(enabled = FALSE)) |>
  hc_xAxis(max = mol_q99, title = list(text = "depth")) |>
  hc_yAxis(title = list(text = "% intervals"))  |>
  hc_caption(text = "inferred molecules calculate \"effective\" or \"molecule\" coverage") |>
  hc_title(text = "Distribution of Alignment Depths")  |>
  hc_exporting(enabled = T, filename = paste0(samplename, ".cov"),
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

## Sumheader
### sumhead {.no-title}
<h2> Coverage Summary Information </h2>
These tables will help you understand the sequence coverage of the sample.

## Tableheaders
### Sumdesc {.no-title}
The table below shows the global and per-contig average depth and standard 
deviation per `r windowskb`kbp intervals **including** intervals whose depth is flagged
an outlier in the data. 

### outlierdesc {.no-title}
The table below shows the `r windowskb`kbp intervals considered outliers, as determined by 
having depth greater than the 99th percentile (`r q99`) of aligment depths. 

## Summary information 
### Averages
```{r depth_table}
DT::datatable(
  contig_avg, 
  rownames = F, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Brtip',
    scrollX = TRUE,
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_align_depth_avg")))
  ),
  colnames = c('Contig', 'Average Depth', 'Standard Deviation', 'Average Depth (molecules)', 'Standard Deviation (Molecules)'),
  fillContainer = T,
  height = "fit-content"
)
```

### Outliers
```{r depth_plot}
DT::datatable(
  outliers, 
  rownames = F, 
  extensions = 'Buttons', 
  options = list(
    dom = 'Brtip',
    scrollX = TRUE,
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_align_depth_outlier_avg")))
  ),
  colnames = c('Contig', 'Position', 'Depth', 'Depth (molecules)'),
  fillContainer = T
)
```

## Circleplot {.no-title data-height=900}
### pltdsc {.no-title data-width=150}
```{r disttext}
if(snakemake@params$mol_dist > 0){
  dist_text <- paste0("from the linked-read information with a distance threshold of ", snakemake@params$mol_dist)
} else {
  dist_text <- "by the internal EMA aligner heuristics"
}
```

<h2> Depth and Coverage Across the Genome </h2>
To the right is a circular plot summarizing the depth information across up to 30 of the largest contigs.
For clarity, this visualization truncates coverage at the 99th percentile (`r q99` for alignments and `r mol_q99` for molecule coverage).
If you are unfamiliar with this kind of visualization, it's a circular version of a linear genome. 
Each arc (segment) is a different contig, from position 0 to the end of the contig, and is labelled by the contig name.
The internal (grey) rings are a barplot where each bar represents the alignment depth at a `r windowskb` kilobase
genomic interval. The ring (purple bars) is the number of reads that had a _proper_ alignment in the `r windowskb` kilobase interval, where
"proper" refers to a read not marked as a duplicate or flagged with the SAM `UNMAP`, `SECONDARY`,  or `QCFAIL` flags. The outer ring (green bars),
is the number of _molecules_ spanning that interval, where molecules are inferred `r dist_text`.

You may hover your cursor over variants to view their positions, pan by clicking and dragging,
and zoom using scroll (mouse or touchpad). In case you become unable to scroll up from the plot due to these interactive 
features, place your cursor over this left column or the navigation bar of this report and you will
be able to scroll the report instead of zooming on the plot.

### Coverage across the genome
```{r limit_contigs}
# Find the 30 largest contigs
contigs <- group_by(tb, contig) %>%
  summarize(size = max(position)) %>%
  arrange(desc(size))

# limit the data to only the 30 largest contigs
if (nrow(contigs) > 30){
    .contigs <- contigs[1:30, ]
    tb <- filter(tb, contig %in% .contigs$contig)
} else {
    .contigs <- contigs
}
```

```{r circleplot, fig.align='center', out.width= "80%", out.height="900px"}
genomeChr <- .contigs$size
names(genomeChr) <- .contigs$contig
genomeChr <- as.list(genomeChr)
inner <- c(0.2, 0.6)
outer <- c(0.61, 0.97)
tracks <- BioCircosTracklist()

# Add one track for each chromosome
for (i in names(genomeChr)){
  # Define histogram/bars to be displayed
  chrcov <- tb[tb$contig == i,]
  #chrcov$depth
  # Add a track with bars on the i-th chromosome
  tracks = tracks + BioCircosBarTrack(
    paste0("cov_", i),
    chromosome = i, 
    starts = chrcov$position - windowsize, ends = chrcov$position,
    values = pmin(chrcov$depth, q99),
    color = "#56486d",
    minRadius = inner[1],
    maxRadius = inner[2]
  )
  tracks = tracks + BioCircosBarTrack(
    paste0("molcov_", i),
    chromosome = i, 
    starts = chrcov$position - windowsize, ends = chrcov$position,
    values = pmin(chrcov$mol_depth, mol_q99),
    color = "#699867",
    minRadius = outer[1],
    maxRadius = outer[2]
  )
}
# Add background
tracks = tracks + BioCircosBackgroundTrack(
  "depth_background",
  fillColors = "#f3f3f3",
  borderColors = "#C9C9C9",
  minRadius = inner[1],
  maxRadius = inner[2]
)
tracks = tracks + BioCircosBackgroundTrack(
  "molcov_background",
  fillColors = "#f3f3f3",
  borderColors = "#C9C9C9",
  minRadius = outer[1],
  maxRadius = outer[2]
  )
BioCircos(
  tracks,
  displayGenomeBorder = F,
  genome = genomeChr,
  chrPad = 0.02,
  genomeTicksDisplay = F,
  BARMouseOverColor = "#1cff42",
  BARMouseOverTooltipsHtml05 = "Mean Depth: ",
  genomeLabelDy = 0,
  width = "100%"
)
```