---
title: "Alignment Barcode Summary"
params:
  indir: 'reports/data/bxstats/'
---
`r format(Sys.time(), '🗓️ %d %B, %Y 🕔 %H:%M')`

```{r imports, results = F}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("dplyr","tidyr", "highcharter","DT")
```

```{r nxx_and_process_funs, results = F}
NX <- function(lengths, X=50){
  lengths <- as.numeric(sort(lengths, decreasing=TRUE))
  index <- which(cumsum(lengths) / sum(lengths) >= X/100)[1]
  return(lengths[index])
}

avg_percent_of <- function(target, other){
  # omit zeroes, take the means of target and other, then get a percent
  x <- target[target != 0]
  y <- other[other != 0]
  return(
    round(mean(x/(x+y)) * 100, 2)
  )
}

std_error <- function(x){
  sd(x)/sqrt(length((x)))
}

process_input <- function(infile){
  samplename <- gsub(".bxstats.gz", "", basename(infile))
  tb <- read.table(infile, header = T, sep = "\t") %>% select(-4, -5)
  tb$valid <- tb$molecule
  tb[tb$valid != "invalidBX", "valid"] <- "validBX"
  tb$valid <- gsub("BX", " BX", tb$valid)
  # isolate non-singletons b/c molecules with 1 read pair aren't linked reads
  multiplex_df <- filter(tb, valid == "valid BX", reads >= 2)
  singletons <- sum(tb$reads < 2 & tb$valid == "valid BX")
  tot_uniq_bx <- read.table(infile, header = F, sep = "\n", as.is = T, skip = nrow(tb) + 1, comment.char = "+")
  tot_uniq_bx <- gsub("#total unique barcodes: ", "", tot_uniq_bx$V1[1]) |> as.integer()
  tot_mol <- sum(tb$valid == "valid BX")
  tot_valid_reads <- sum(tb[tb$valid == "valid BX", "reads"])
  avg_reads_per_mol <- round(mean(multiplex_df$reads),1)
  sem_reads_per_mol <- round(std_error(multiplex_df$reads), 2)
  tot_invalid_reads <- sum(tb[tb$valid == "invalid BX", "reads"])
  avg_mol_cov <- round(mean(multiplex_df$coverage_inserts), 2)
  sem_mol_cov <- round(std_error(multiplex_df$coverage_inserts), 4)
  avg_mol_cov_bp <- round(mean(multiplex_df$coverage_bp), 2)
  sem_mol_cov_bp<- round(std_error(multiplex_df$coverage_bp), 4)
  n50 <- NX(multiplex_df$length_inferred, 50)
  n75 <- NX(multiplex_df$length_inferred, 75)
  n90 <- NX(multiplex_df$length_inferred, 90)
  datarow <- data.frame(
    sample = samplename,
    totalreads = tot_valid_reads + tot_invalid_reads,
    totaluniquemol = tot_mol,
    singletons = singletons,
    nonsingletons = nrow(multiplex_df),
    totaluniquebx = tot_uniq_bx,
    molecule2bxratio = round(tot_mol / tot_uniq_bx,2),
    totalvalidreads = tot_valid_reads,
    totalinvalidreads = tot_invalid_reads,
    totalvalidpercent = round(tot_valid_reads/(tot_valid_reads + tot_invalid_reads) * 100,2),
    averagereadspermol = avg_reads_per_mol,
    sereadspermol = sem_reads_per_mol,
    averagemolcov = avg_mol_cov,
    semolcov = sem_mol_cov,
    averagemolcovbp = avg_mol_cov_bp,
    semolcovbp = sem_mol_cov_bp,
    N50 = n50,
    N75 = n75,
    N90 = n90
  )
  return(datarow)
}

```

```{r setup_df}
infiles <- list.files(params$indir, "bxstats.gz", full.names = TRUE)
aggregate_df <- Reduce(rbind, Map(process_input, infiles))
aggregate_df[is.na(aggregate_df)] <- 0

if(nrow(aggregate_df) == 0){
  print("All input files were empty")
  knitr::knit_exit()
}
```

# General
## General Barcode Information from Alignments
<h2> General Barcode Information from Alignments </h2>
This report aggregates the barcode-specific information from the alignments
that were created using `harpy align`. Detailed information for any one sample
can be found in that sample's individual report. The table below is an aggregation
of data for each sample based on their `*.bxstats.gz` file. Every column after `% valid bx`
ignores singletons in its calculations. Hover the column name to display a tooltip with the column's description.

```{r}
#| content: valuebox
#| title: "Samples"
list(
  color = "light",
  value = nrow(aggregate_df)
)
```

```{r}
#| content: valuebox
#| title: "Avg % Nonsingletons"
list(
  color = "light",
  value = avg_percent_of(aggregate_df$nonsingletons, aggregate_df$singletons)
)
```

```{r}
#| content: valuebox
#| title: "Avg Percent Valid Bx"
list(
  color = "light",
  value = round(mean(aggregate_df$totalvalidpercent[aggregate_df$totalvalidpercent != 0]), 2)
)
```

```{r}
#| content: valuebox
#| title: "Avg Reads/Molecule"
list(
  color = "light",
  value = round(mean(aggregate_df$averagereadspermol[aggregate_df$averagereadspermol != 0]), 2)
)
```

```{r}
#| content: valuebox
#| title: "N50"
list(
  color = "light",
  value = round(mean(aggregate_df$N50[aggregate_df$N50 != 0]), 0)
)
```

```{r}
#| content: valuebox
#| title: "N75"
list(
  color = "light",
  value = round(mean(aggregate_df$N75[aggregate_df$N75 != 0]), 0)
)
```

```{r}
#| content: valuebox
#| title: "N90"
list(
  color = "light",
  value = round(mean(aggregate_df$N90[aggregate_df$N90 != 0]), 0)
)
```

## Sampletable

```{r}
column_description <- c(
  "name of the sample",
  "total number of alignments",
  "the unique DNA molecules as inferred from linked-read barcodes",
  "molecules composed of one single-end or paired-end sequence, effectively not a linked read",
  "molecules composed of more than two single-end or paired-end sequences, in other words, molecules with linked-read information",
  "number of unique barcodes, which may differ from unique molecules after deconvolution",
  "molecule-to-barcode ratio, which helps diagnose deconvolution performance",
  "alignments with a valid AXXCXXBXXDXX barcode, where XX is not 00",
  "alignments with a invalid 00 barcode segment",
  "percent of valid-to-invalid barcoded alignments",
  "average number of reads per unique molecule",
  "standard error of the mean number of reads per unique molecule",
  "average percent molecule alignment coverage, where coverage includes unsequenced gaps between linked reads",
  "standard error of the mean percent molecule alignment coverage",
  "average molecule alignment coverage in base pairs, where coverage includes unsequenced gaps between linked reads",
  "standard error of the mean molecule alignment base-pair coverage",
  "N50 of inferred molecules",
  "N75 of inferred molecules",
  "N90 of inferred molecules"
  )

headerCallback <- c(
  "function(thead, data, start, end, display){",
  paste0(" var tooltips = ['", paste(column_description, collapse = "','"), "'];"),
  "  for(var i=0; i<tooltips.length; i++){",
  "    $('th:eq('+i+')',thead).attr('title', tooltips[i]);",
  "  }",
  "}"
)

DT::datatable(
  aggregate_df,
  rownames = F,
  extensions = 'Buttons', 
  caption = 'Per-Sample Alignment Information Pertaining to Barcodes',
  fillContainer = T,
  colnames = c("sample", "alignments", "unique molecules", "singletons", "nonsingletons", "unique barcodes", "molecule:bx ratio", "valid bx alignments", "invalid bx alignments", "% valid bx", "avg reads/molecule", "SE reads/molecule", "avg molecule coverage %","SE molecule coverage %", "avg molecule coverage (bp)", "SE molecule coverage (bp)", "N50", "N75","N90"),
  options = list(
    dom = 'Brtp',
    scrollX = TRUE,
    buttons = list(list(extend = "csv",filename = "per_sample_barcode_alignment")),
    headerCallback = JS(headerCallback)
  )
)
```

```{r sampleplot_height}
n_samples <- nrow(aggregate_df)
figheight <- 3 + (0.4 * n_samples)
```

## NX plot
::: {.card title="NX Information"}

The **NX** metric (e.g. **N50**) is the length of the shortest molecule in the group of longest molecules that together
represent at least **X%** of the total molecules by length. For example, `N50` would be the shortest molecule in the 
group of longest molecules that together represent **50%** of the total molecules by length (sort of like a median).
These are the distributions of three common NX metrics (N50, N75, N90) across all samples.

```{r nxxplot}
if(length(unique(aggregate_df$N50)) > 1 && length(unique(aggregate_df$N75)) > 1 && length(unique(aggregate_df$N90)) > 1){
  highchart() |>
    hc_chart(type = "area", animation = F) |>
    hc_add_series(data = density(aggregate_df$N50), name = "N50", type = "areaspline") |>
    hc_add_series(data = density(aggregate_df$N75), name = "N75", type = "areaspline") |>
    hc_add_series(data = density(aggregate_df$N90), name = "N90", color = "#d3805f", type = "areaspline") |>
    hc_tooltip(enabled = FALSE) |>
    hc_caption(text = "Values derived using non-singleton molecules") |>
    hc_title(text = "NX Stats Across Samples") |>
    hc_xAxis(title = list(text = "length (bp)"), min = 0) |>
    hc_yAxis(title = list(text = "density")) |>
    hc_exporting(enabled = T, filename = "NX.stats",
      buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
    )
} else if (nrow(aggregate_df) == 1){
  nx_df <- data.frame("NX" = c("N50", "N75", "N90"), "value" = c(aggregate_df$N50, aggregate_df$N75, aggregate_df$N90))
  hchart(nx_df, animation = FALSE, type = "column", hcaes(x = NX, y = value, color = NX), name = "value") |>
    hc_tooltip(enabled = TRUE) |>
    hc_caption(text = "Values derived using non-singleton molecules") |>
    hc_title(text = "NX Stats") |>
    hc_xAxis(title = list(text = "Stat"), categories = nx_df$NX) |>
    hc_yAxis(title = list(text = "length (bp)")) |>
    hc_exporting(enabled = T, filename = "NX.stats",
      buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
    )
} else {
  cat("One of more of the NXX stats don't have enough unique values (minimum: 2) to plot distributions:\n")
  cat("  N50:", unique(aggregate_df$N50), "\n")
  cat("  N75:", unique(aggregate_df$N75), "\n")
  cat("  N90:", unique(aggregate_df$N90), "\n")
}
```
:::

### Reads per molecule
::: {.card title="Reads per molecule"}
This distribution shows the average number of reads per molecule across your samples.

```{r molper_dist}
if(length(unique(aggregate_df$averagereadspermol)) > 1){
  hchart(density(aggregate_df$averagereadspermol), type = "areaspline", color = "#e1a42b", name = "Reads per Molecule", animation = F, marker = list(enabled = FALSE)) |>
    hc_title(text = "Average Reads Per Molecule") |>
    hc_xAxis(title = list(text = "reads per molecule (average)")) |>
    hc_yAxis(title = list(text = "density")) |>
    hc_tooltip(enabled = FALSE) |>
    hc_exporting(enabled = T, filename = "reads_per_molecule",
      buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
    )
} else {
  cat("Not enough unique values for reads per molecule to plot distributions (minimum: 2).\n")
}
```
:::

## Distribution of valid bx alignments
::: {.card title="Distribution of alignments with valid barcodes"}
This is a distribution of what percent of total alignments each sample
had valid haplotag barcodes (`AXXCXXBXXDXX` where `XX` is not `00`).

```{r perc_valid_dist}
hs <- hist(
  aggregate_df$totalvalidpercent,
  breaks = seq(0,100, by = 5),
  plot = F
)
hs$counts <- round(hs$counts / sum(hs$counts) * 100, 4)
hs <- data.frame(val = hs$breaks[-1], freq = hs$counts)

hchart(hs, "areaspline", hcaes(x = val, y = freq), color = "#8484bd", name = "Percent Alignments", marker = list(enabled = FALSE), animation = F) |>
  hc_title(text = "Percent of Alignments with Valid BX tags") |>
  hc_xAxis(min = 0, max = 100, title = list(text = "% alignments with valid BX tag")) |>
  hc_yAxis(max = 100, title = list(text = "% samples")) |>
  hc_tooltip(crosshairs = TRUE) |>
  hc_exporting(enabled = T, filename = "percent.valid.dist",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```
:::

### Linked-read library performance
::: {.card title="Linked-read library performance"}
This scatterplot is a diagnostic that shows the relationship between the proportion of nonsingleton
reads (reads with linked info) compared to total reads. Each point is a sample and is colored by the
mean number of reads per molecule for that sample.

```{r lr_ratio}
hchart(aggregate_df, "scatter", hcaes(x = totalreads, y = round(nonsingletons / (nonsingletons + singletons), 2), color = averagereadspermol, name = sample, molper = averagereadspermol), name = "Linked-Read Performance", animation = F, marker = list(enabled = FALSE)) |>
  hc_title(text = "Linked Reads vs Total Reads") |>
  hc_xAxis(title = list(text = "total reads")) |>
  hc_yAxis(title = list(text = "proportion of reads with linked-read information")) |>
  hc_tooltip(
    crosshairs = FALSE,
    pointFormat= '<b>{point.name}</b><br>reads: <b>{point.x}</b><br>% linked reads: <b>{point.y}</b><br>reads per molecule: <b>{point.molper}</b></br>'
    ) |>
  hc_exporting(enabled = T, filename = "linkedread.performance",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```
:::

# Per-Sample
<h2> Per-Sample Metrics </h2>
Below is a series of plots showing per-sample metrics. The meaning of percentages and error bars
are provided in the bottom-left captions of the plots.

## Per-sample plots
### percent valid
```{r perc_valid_per_sample, fig.height=figheight}
hchart(aggregate_df, "xrange", animation = F, groupPadding = 0.0001,
  hcaes(x = 0, x2 = totalreads, y = rev(0:(n_samples-1)), partialFill = totalvalidpercent/100), dataLabels = list(enabled = TRUE)) |> 
  hc_xAxis(title = list(text = "Total Alignments")) |> 
  hc_yAxis(title = FALSE, gridLineWidth = 0, categories = rev(aggregate_df$sample)) |>
  hc_caption(text = "Percentage represents the percent of alignments with a valid BX barcode, shown in dark green. Light green represents invalid BX barcodes.") |>
  hc_tooltip(enabled = FALSE) |>
  hc_colors(color = "#95d8a7") |>
  hc_title(text = "Percent Alignments with Valid BX Barcode") |>
  hc_exporting(enabled = T, filename = "BX.valid.alignments",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

### nonsingletons percent
```{r perc_nonsingleton, fig.height=figheight}
hchart(aggregate_df, "xrange", animation = F, groupPadding = 0.0001,
  hcaes(x = 0, x2 = nonsingletons + singletons, y = rev(0:(n_samples-1)), partialFill = round(nonsingletons / (nonsingletons + singletons), 4)), dataLabels = list(enabled = TRUE)) |> 
  hc_xAxis(title = list(text = "Total Alignments with Valid Barcodes")) |> 
  hc_yAxis(title = FALSE, gridLineWidth = 0, categories = rev(aggregate_df$sample)) |>
  hc_caption(text = "Percentage represents the percent of molecules composed of more than 2 single-end or paired-end sequences, shown in dark blue. Light blue represents the singletons.") |>
  hc_tooltip(enabled = FALSE) |>
  hc_colors(color = "#8dc6f5") |>
  hc_title(text = "Percent Non-Singleton Molecules") |>
  hc_exporting(enabled = T, filename = "BX.nonsingletons",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

### reads per molecule
```{r reads_per_mol_sample, fig.height=figheight}
err_df <- data.frame(x = 0:(n_samples-1), y = aggregate_df$averagereadspermol, low = aggregate_df$averagereadspermol - aggregate_df$sereadspermol, high = aggregate_df$averagereadspermol + aggregate_df$sereadspermol)
highchart() |>
  hc_chart(inverted=TRUE, animation = F, pointPadding = 0.0001, groupPadding = 0.0001) |>
  hc_add_series(data = aggregate_df, type = "scatter", name = "mean", hcaes(x = 0:(nrow(aggregate_df)-1), y = averagereadspermol), marker = list(radius = 8), color = "#6d73c2", zIndex = 1) |>
  hc_add_series(data = err_df, type = "errorbar", name = "standard error", linkedTo = ":previous", zIndex = 0, stemColor = "#8186c7", whiskerColor = "#8186c7") |>
  hc_xAxis(title = FALSE, gridLineWidth = 0, categories = aggregate_df$sample) |>
  hc_title(text = "Average Reads Per Molecule") |>
  hc_caption(text = "Error bars show the standard error of the mean of non-singleton molecules") |>
  hc_tooltip(crosshairs = TRUE, pointFormat= '<b>{point.y}</b>') |>
  hc_exporting(enabled = T, filename = "avg.reads.per.molecule",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

### average molecule coverage
```{r avg_mol_cov, fig.height=figheight}
err_df <- data.frame(x = 0:(n_samples-1), y = aggregate_df$averagemolcov, low = aggregate_df$averagemolcov - aggregate_df$semolcov, high = aggregate_df$averagemolcov + aggregate_df$semolcov)
err_df_bp <- data.frame(x = 0:(n_samples-1), y = aggregate_df$averagemolcovbp, low = aggregate_df$averagemolcovbp - aggregate_df$semolcovbp, high = aggregate_df$averagemolcovbp + aggregate_df$semolcovbp)

highchart() |>
  hc_chart(inverted=TRUE, animation = F, pointPadding = 0.0001, groupPadding = 0.0001) |>
  hc_add_series(data = aggregate_df, type = "scatter", name = "Inferred Fragments", hcaes(x = 0:(nrow(aggregate_df)-1), y = averagemolcov), marker = list(radius = 8), color = "#df77b5", zIndex = 1) |>
  hc_add_series(data = err_df, type = "errorbar", name = "standard error", linkedTo = ":previous", zIndex = 0, stemColor = "#ef94ca", whiskerColor = "#ef94ca") |>
  hc_add_series(data = aggregate_df, type = "scatter", name = "Aligned Base Pairs", hcaes(x = 0:(nrow(aggregate_df)-1), y = averagemolcovbp), marker = list(radius = 8), color = "#924596", zIndex = 1) |>
  hc_add_series(data = err_df_bp, type = "errorbar", name = "standard error", linkedTo = ":previous", zIndex = 0, stemColor = "#916094", whiskerColor = "#916094") |>
  hc_xAxis(title = FALSE, gridLineWidth = 0, categories = aggregate_df$sample) |>
  hc_title(text = "Average Molecule Percent Coverage") |>
  hc_caption(text = "Error bars show the standard error of the mean percent coverage of non-singleton molecules") |>
  hc_tooltip(crosshairs = TRUE, pointFormat= '<b>{point.y}</b>') |>
  hc_exporting(enabled = T, filename = "avg.molecule.coverage",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```
