---
title: "Summary - `r params$vcf`"
params:
  infile: 'variants.raw.stats'
  vcf: 'variants.raw'
---
`r format(Sys.time(), '🗓️ %d %B, %Y 🕔 %H:%M')`

```{r package_imports, results = F}
library(magrittr)
library(tidyr)
library(DT)
library(highcharter)
library(scales)
```

```{r}
statsfile <- params$infile
dataL <- readLines(statsfile)
bcf <- gsub(".stats$", ".bcf", statsfile)
```
# SNPs
##
```{r General Stats}
.snL <- grepl("^SN", dataL)
sn <- read.table(text=dataL[.snL], sep = "\t")[,3:4]
names(sn) <- c("Metric", "Number")
sn$Metric <- gsub("number of ", "", sn$Metric)
rownames(sn) <- sn$Metric
sn <- as.data.frame(t(sn[2]))
```
```{r}
#| content: valuebox
#| title: "Samples"
list(
  color = "light",
  value = prettyNum(sn[1,1], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "Records"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,2], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "SNPs"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,4], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "MNPs"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,5], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "Indels"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,6], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "Multiallelic Sites"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,8], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "Multiallelic SNPs"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,9], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "No-Alts"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,3], big.mark = ",")
)
```
```{r}
#| content: valuebox
#| title: "Others"
list(
  color = "#dfdfdf",
  value = prettyNum(sn[1,7], big.mark = ",")
)
```

## Quality Statistics
```{r qual data}
.qual <- grepl("^QUAL", dataL)
if(sum(.qual) == 0){
  do_qual <- FALSE
  cat(paste0("No QUAL section in ", statsfile, " found\n"))
} else {
  do_qual <- TRUE
}
if(do_qual){
  qual <- read.table(text=dataL[.qual])[ , c(3,4,7)]
  names(qual) <- c("QualityScore", "SNPs", "Indels")
  qual[names(qual)] <- sapply(qual[names(qual)],as.numeric)
}
```
```{r resizehist}
# function to merge QUAL bins into a width of 10
mergebins <- function(x){
  bins <- seq(0, max(x[,1]), 10)
  lastbin <- max(x[,1])
  bins <- c(bins, lastbin)
  dict <- vector(mode="list", length=length(bins))
  names(dict) <- as.character(bins)
  for(i in 1:nrow(x)){
    databin <- x[i,1]
    key <- max(which(databin >= bins))
    dict[[key]] <- c(dict[[key]], x[i,2])
  }
  dict <- lapply(dict, sum)
  df <- as.data.frame(t(as.data.frame(dict)))
  df$V2 <- as.integer(gsub("X", "", row.names(df)))

  df <- df[,c(2,1)]
  names(df) <- names(x)
  row.names(df) <- NULL
  return(df)
}

```

```{r hist_function}
hist_quantile <- function(x, q){
  # find the qth percentile of the histogram x
  # Calculate the cumulative sum of the counts
  cumulative_counts <- cumsum(x[,2])
  # Normalize to create a cumulative distribution function (CDF)
  cdf <- cumulative_counts / max(cumulative_counts)
  # Find the bin that is closest to the qth percentile
  index <- which.min(abs(cdf - q))
  # The qth percentile value is the right edge of this bin
  # cap it at the last index if it exceeds
  index <- min(nrow(x), index + 1)
  return(x[index,1])
}
```
##
::: {.card title="SNP Genotype Quality"}
These are per-locus SNP quality statistics, which correspond to the Quality (`QUAL`) values of `bcftools stats`. 
Each quality score is has a semi-open bin width of 10, where `0` would be considered "a QUAL score
greater than or equal to `0` and less than `10`". The histograms are truncated at the 99th percentile
for visual clarity.

```{r snpqual plot}
if(do_qual){
  rebinned_qual <- mergebins(qual[,-3])
  q99 <- hist_quantile(rebinned_qual, 0.99)
  rebinned_qual <- rebinned_qual[rebinned_qual$QualityScore < q99, ]
  hchart(rebinned_qual, "areaspline", hcaes(x = QualityScore, y = SNPs), color = "#8fbbbc", name = "count", marker = list(enabled = FALSE)) |>
    hc_title(text = "SNP Genotype Quality") |>
    hc_subtitle(text = "Shown in bins of width 10") |>
    hc_caption(text = paste("99th percentile:", q99)) |>
    hc_xAxis(title = list(text = "quality score"), type = "logarithmic") |>
    hc_yAxis(title = list(text = "count")) |>
    hc_tooltip(crosshairs = TRUE,
      formatter = JS("function () {return '<b>Quality ' + this.x + '</b><br>Count: <b>' + this.y + '</b>';}") 
    ) |>
    hc_exporting(enabled = T, filename = "snp.quality",
      buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
    )
} else {
  cat(paste0("No QUAL section in ", statsfile, " found\n"))
}
```
:::

```{r findzero_function}
# find index when the values are consistently 0 to hopefully remove an uninformative tail from the plot
find_zero_runs <- function(vec) {
  rle_result <- rle(vec)
  zero_runs <- which(rle_result$values == 0 & rle_result$lengths >= 20)
  if (length(zero_runs) > 0) {
    return(sum(rle_result$lengths[1:(zero_runs[1]-1)]) + 1)
  } else {
    return(NA)
  }
}
```
### {.tabset}
::: {.card title="Depth per Genotype"}
These are per-locus statistics, which correspond to the Depth (`DP`) calculations of `bcftools stats`

```{r depth plot, out.width="100%"}
#| title: Depth per Genotype

.dpL <- grepl("^DP", dataL)
if(sum(.dpL) == 0){
  cat(paste0("No DP section in ", statsfile, " found\n"))
} else {
  dp <- read.table(text=dataL[.dpL])[ , 3:7]
  names(dp) <- c("Bin", "Genotypes", "PercentGenotypes", "NumberSites", "PercentSites")
  dp$PercentGenotypes <- round(dp$PercentGenotypes, digits = 6)
  not_low <- dp$PercentGenotypes > 0.01
  dp$PercentGenotypes[not_low] <- round(dp$PercentGenotypes[not_low], digits = 2)
  dp$PercentSites <- round(dp$PercentSites, digits = 4)
  # cut off at trailing zeroes
  cutoff <- find_zero_runs(dp$PercentGenotypes) - 1
  if(is.na(cutoff)){
    cutoff <- nrow(dp)
  }
  hchart(dp[1:cutoff,], "areaspline", hcaes(x = Bin, y = PercentGenotypes), name = "percent", color = "#F28500", marker = list(enabled = FALSE)) |>
    hc_title(text = "Depth per Genotype") |>
    hc_xAxis(title = list(text = "depth bin"), type = "logarithmic") |>
    hc_yAxis(title = list(text = "% of genotypes")) |>
    hc_tooltip(crosshairs = TRUE,
      formatter = JS("function () {return '<b>depth: ' + this.x + '</b><br>Percent: <b>' + this.y + '</b>';}") 
    ) |>
    hc_exporting(enabled = T, filename = "genotype.depth",
      buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
}
```
:::

```{r depth table, out.width="100%"}
#| title: Table

if(sum(.dpL) == 0){
  cat(paste0("No DP section in ", statsfile, " found\n"))
} else {
  column_description <- c(
    "depth bin as computed by bcftools",
    "number of genotypes in the bin",
    "percent of genotypes in the bin",
    "number of sites in the bin",
    "percent of sites in the bin"
    )

  headerCallback <- c(
    "function(thead, data, start, end, display){",
    paste0(" var tooltips = ['", paste(column_description, collapse = "','"), "'];"),
    "  for(var i=0; i<tooltips.length; i++){",
    "    $('th:eq('+i+')',thead).attr('title', tooltips[i]);",
    "  }",
    "}"
  )

  DT::datatable(
    dp,
    rownames = F,
    filter = "top",
    extensions = 'Buttons',
    colnames = c("Depth Bin", "Genotypes", "% Genotypes", "Sites", "% Sites"),
    options = list(
      dom = 'Brtp',
      buttons = list(list(extend = "csv", filename = "align_depth_bins")),
      paging = T,
      headerCallback = JS(headerCallback)
    )
  )
}
```


# Indels
##
::: {.card title="Insertion and Deletion Distribution"}
This is the distribution of insertions and deletions based on length and frequency. The deletion
length is presented here as negative values for visual clarity.

```{r Indel Stats, paged.print=TRUE, out.width="100%"}
.iddL <- grepl("^IDD", dataL)
if(sum(.iddL) == 0){
  cat(paste0("No IDD section in ", statsfile, " found, suggesting no indels are present\n"))
} else {
  idd <- read.table(text=dataL[.iddL])[ ,3:4]
  names(idd) <- c("Length", "Count")
  idd$Type <- idd$Length > 0
  idd$Type <-  gsub(TRUE, "Insertion", idd$Type)
  idd$Type <-  gsub(FALSE, "Deletion", idd$Type)
  if(nrow(idd) > 1){
    hchart(idd, "areaspline", hcaes(x = Length, y = Count, group = Type, marker = list(enabled = FALSE))) |>
      hc_colors(c("#e6a037","#6daace")) |>
      hc_title(text = "Insertion-Deletion Distribution") |>
      hc_xAxis(
        title = list(text = "indel length"),
        labels = list(formatter = JS("function () {return Math.abs(this.value);}"))
      ) |>
      hc_yAxis(title = list(text = "number of sites")) |>
      hc_tooltip(crosshairs = TRUE,
        formatter = JS("function () {return '<b>' + this.series.name + '</b><br>Length: <b>' + Math.abs(this.x) + '</b><br/>Count: <b>' + this.y + '</b>';}") 
  ) |>
      hc_exporting(enabled = T, filename = "indel.distribution",
        buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
      )
  } else {
    cat("At least two indels are required to plot a distribution\n")
  }
}
```
:::

### Indel Quality
::: {.card title="Indel Genotype Quality"}
These are per-locus indel quality statistics, which correspond to the Quality (`QUAL`) values of `bcftools stats`. 
Each quality score is has a semi-open bin width of 10, where `0` would be considered "a QUAL score
greater than or equal to `0` and less than `10`". The histograms are truncated at the 99th percentile
for visual clarity.
```{r indelqual plot}
if(do_qual){
  rebinned_qual <- mergebins(qual[,-2])
  q99 <- hist_quantile(rebinned_qual, 0.99)
  rebinned_qual <- rebinned_qual[rebinned_qual$QualityScore < q99, ]
  if(nrow(rebinned_qual) > 1){
    hchart(rebinned_qual, "areaspline", hcaes(x = QualityScore, y = Indels), name = "count", color = "#9393d2", marker = list(enabled = FALSE)) |>
      hc_title(text = "Indel Genotype Quality") |>
      hc_subtitle(text = "Shown in bins of width 10") |>
      hc_caption(text = paste("99th percentile:", q99)) |>
      hc_xAxis(title = list(text = "quality score"), type = "logarithmic") |>
      hc_yAxis(title = list(text = "count")) |>
      hc_tooltip(crosshairs = TRUE,
        formatter = JS("function () {return '<b>Quality ' + this.x + '</b><br>Count: <b>' + this.y + '</b>';}") 
      ) |>
      hc_exporting(enabled = T, filename = "indel.quality",
        buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
      )
  } else {
    cat("At least two quality values are required to plot a distribution\n")
  }
} else {
  cat(paste0("No QUAL section in ", statsfile, " found\n"))
}
```
:::

# Per-Sample
##

```{r read_persample}
.pscL <- grepl("^PSC", dataL)
psc <- read.table(text=dataL[.pscL])[ ,3:14]
names(psc) <- c("Sample", "HomRef", "HomAlt", "Het", "Transitions", "Transversions", "Indels",	"MeanDepth", "Singletons",	"HapRef", "HapAlt", "Missing")
tidy_psc <- pivot_longer(psc[,c(1,2,3,4,5,6,7,9,12)], -Sample , names_to = "Metric", values_to = "Count")
tidy_psc$Sample <- as.factor(basename(tidy_psc$Sample))
figheight <- 3 + (0.6 * nrow(psc))
```

::: {.card title="Individual Stats (Plot)"}
These are per-locus statistics, which correspond to the Per-Sample Counts (`PSC`) calculations of `bcftools stats`. The table reflects the data visualized in the plot tab. `HomAlt` and `HomRef` refer to homozygous for the Reference and Alternative alleles, respectively.

```{r Individual plot, fig.height=figheight}
#| title: Individual Stats (plot)
highchart() |>
  hc_chart(inverted=TRUE, animation = F) |>
  hc_add_series(data = tidy_psc, type = "scatter", hcaes(y = Count, x = Sample, group = Metric)) |>
  hc_colors(c("#41739c", "#eb7e2b", "#aa042e", "#837fbd", "#202020", "#9dc96c", "#E6AB02", "#d39ab7")) |>
  hc_title(text = "Individual Statistics") |>
  hc_xAxis(type = "category", title = list(text = ""), categories = levels(tidy_psc$Sample)) |>
  hc_yAxis(title = list(text = "count")) |>
  hc_tooltip(
    crosshairs = TRUE, animation = F,
    pointFormat= '<b>{point.category}</b><br>Count: <b>{point.y:.0f}</b>'
  ) |>
  hc_exporting(enabled = T, filename = "samples.variantstats",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

:::

```{r Individual table}
#| title: Individual Stats (Table)
#| expandable: true
DT::datatable(
  psc, rownames = F,
  fillContainer = T,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = "snp_indiv_stats")),
    scrollX = TRUE,
    paging = F
  )
)
```

