---
title: "LEVIATHAN SV - `r params$sample`"
brand:
  logo: 
    medium: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_navbar.png
execute:
    echo: false
    warning: false
embed-resources: true
format: 
  dashboard:
    theme: materia
    scrolling: true
    expandable: false
    logo: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_navbar.png
    nav-buttons:
      - icon: book-half
        href: https://pdimens.github.io/harpy
      - icon: bug
        href: https://github.com/pdimens/harpy/issues/new/choose
      - icon: code-slash
        href: https://github.com/pdimens/harpy
params:
  faidx: 'genome.fasta.fai'
  statsfile: 'sample.stats'
  sample: 'sample'
  contigs: 'default'
---
`r format(Sys.time(), '🗓️ %d %B, %Y 🕔 %H:%M')`

```{r package_imports}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("dplyr","ggplot2","DT","BioCircos", "htmltools")
```

# General Stats
```{r read_inputs}
fai <- params$faidx
statsfile <- params$statsfile
samplename <- params$sample
```

```{r process_statsfile}
sv <- read.table(statsfile, header = T,
  colClasses = c("factor", "factor", "integer", "integer", "character", "factor", "integer", "integer")
)
if (nrow(sv) == 0) {
  cat(paste0("There are no variants in the file ", "`", samplename, ".bcf`"))
  knitr::knit_exit()
}
```
```{r summaryinfo}
summinfo <- sv %>%
  group_by(contig, type) %>%
  summarise(count = n()) %>% 
  group_by(type) %>%
  summarise(
    total = sum(count),
    avg_per_contig = round(mean(count), 2),
    sd = round(sd(count), 2)
)

nvar <- sum(summinfo$total)
```
##
```{r}
#| content: valuebox
#| title: "Variants"
list(
  color = ifelse(nvar > 0, "success", "warning"),
  value = scales::comma(nvar)
)
```
```{r}
#| content: valuebox
#| title: "Breakends"

nbnd <- ifelse("BND" %in% summinfo$type, summinfo$total[which(summinfo$type == "BND")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(nbnd)
)
```
```{r}
#| content: valuebox
#| title: "Deletions"
ndel <- ifelse("DEL" %in% summinfo$type, summinfo$total[which(summinfo$type == "DEL")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ndel)
)
```
```{r}
#| content: valuebox
#| title: "Duplications"
ndup <- ifelse("DUP" %in% summinfo$type, summinfo$total[which(summinfo$type == "DUP")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ndup)
)
```
```{r}
#| content: valuebox
#| title: "Inversions"
ninv <- ifelse("INV" %in% summinfo$type, summinfo$total[which(summinfo$type == "INV")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ninv)
)
```
##
This report details the structural variants identified by [LEVIATHAN](https://github.com/morispi/LEVIATHAN)
[(preprint)](https://doi.org/10.1101/2021.03.25.437002). This tab (`General Stats`) shows overview
information. The `Per-Contig Plot` tab shows you an interactive
plot detailing all variants identified. The variants are given by:

INV (Inversion)
: A segment that broke off and reattached within the same chromosome, but in reverse orientation

DUP (Duplication)
: A type of mutation that involves the production of one or more copies of a gene or region of a chromosome

DEL (Deletion)
: A type of mutation that involves the loss of one or more nucleotides from a segment of DNA

BND (Breakend)
: The endpoint of a structural variant, which can be useful to explain complex variants

##
```{r fix_breakend_length}
sv$length <- gsub("\\.", "1", sv$length) %>%  as.numeric()
```
## Various Stats Tables
::: {.card title="All variants found"}
This table details all the variants LEVIATHAN detected and evidence for those detections.
```{r various stats}
DT::datatable(
  sv,
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtip',
    buttons = list(list(extend = "csv",filename = paste0(samplename , "_sv_leviathan`"))),  
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::
::: {.card title="Variants per Contig"}
This table details each type of structural variant for every contig. 

```{r percontig}
grpstats <- sv %>%
  group_by(contig, type) %>%
  summarise(count = n(), total_bp  = sum(length))

DT::datatable(
  grpstats, 
  rownames = F, 
  filter = "top", 
  extensions = 'Buttons', 
  options = list(
    dom = 'Brtip',
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_sv_leviathan_per_contig"))),
    scrollX = TRUE,
    paging = T
  ),
  fillContainer = T
)
```
:::

# Per-Contig Plot
```{r filter_contigs}
fa.sizes <- read.table(fai, header = F) %>% arrange(desc(2))
colnames(fa.sizes) <- c("contig", "size")

plot_contigs <- params$contigs
if (all(plot_contigs == "default")){
  # make sure that the contigs with SV are the ones being plotted, not the others
  fa.sizes <- fa.sizes[fa.sizes$contig %in% unique(sv$contig), 1:2]
  # limit the data to only the 30 of the largest present contigs
  fa.sizes <- fa.sizes[1:min(nrow(fa.sizes), 30), ]
} else {
  fa.sizes <- fa.sizes[fa.sizes$contig %in% plot_contigs, 1:2]
}

# filter out variants that aren't on one of these contigs
sv <- sv[sv$contig %in% fa.sizes$contig, ]
sv$contig <- factor(sv$contig)
```
##
To the right is a circular plot to visualize the distribution of structural variants
across (up to) 30 of the largest contigs, or whichever contigs were provided. If you are unfamiliar with this kind of visualization,
it's a circular version of a linear genome. Each "wedge" is a different contig, from position 0 to 
the end of the contig, and is labelled by the contig name. Each internal (grey) ring is a plot
of observed structural variants for that contig. The legend shows which colors correspond to which type of variant. To save the plot,
you will need to take a screenshot. Very small variants may be difficult to see or
may not appear on the plot. Deletions are shown as points as they tend to be small
and clump together.

##
::: {.card width=20% title="Legend"}
```{r plot legend}
color.palette <- c("deletion" = "#5a8c84", "duplication" = "#99278a", "inversion" = "#4a9fea", "breakend" = "#c56e34")

data.frame(x = 1, y = 1:4, colour = c("breakend","deletion","duplication", "inversion")) %>% 
ggplot(aes(x, y, fill = colour))+
  geom_point(alpha=0, shape = 22, color = "white")+ # completely transparent rectangular point 
  scale_fill_manual(values=color.palette, drop=FALSE) +
  guides(fill = guide_legend(override.aes = list(alpha=1, size = 28)))+ # showing the point in the legend
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(0.3, 0.5),
        legend.text = element_text(size = 28),
        legend.title=element_text(size=28),
        legend.key = element_rect(fill='NA'),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), 
        plot.background = element_rect(fill='transparent', color=NA),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent', color = "transparent"), 
  ) +
  labs(fill="Variant Type")
```
:::

###
```{r circos}
#| title: "Structural Variants (save by taking a screenshot)"

genomeChr <- fa.sizes$size
names(genomeChr) <- fa.sizes$contig
genomeChr <- as.list(genomeChr)
tracks <- BioCircosTracklist()
radius <- 0.3

# Add one track for each SV type
# inversions
inv <- sv[sv$type == "INV",]
if(nrow(inv) > 0){
  tracks <- tracks + BioCircosArcTrack(
    'Inversions',
    as.character(inv$contig),
    inv$position_start,
    inv$position_end,
    opacities = 0.7,
    values = inv$n_barcodes,
    minRadius = 0.78,
    maxRadius = 0.87,
    colors = color.palette["inversion"],
    labels = "Type: Inversion",
  )
}

# deletions
del <- sv[sv$type == "DEL",]
if(nrow(del) > 0){
  tracks <- tracks + BioCircosSNPTrack(
    'Deletions',
    as.character(del$contig),
    del$position_start,
    values = abs(del$position_end - del$position_start),
    minRadius = 0.63,
    maxRadius = 0.71,
    size = 1,
    shape = "rect",
    opacities = 0.7,
    colors = color.palette["deletion"],
    labels = "Type: Deletion"
  )
}


# duplications
dup <- sv[sv$type == "DUP",]
if(nrow(dup) > 0){
  tracks <- tracks + BioCircosArcTrack(
    'Duplications',
    as.character(dup$contig),
    dup$position_start,
    dup$position_end,
    values = dup$n_barcodes,
    opacities = 0.7,
    minRadius = 0.48,
    maxRadius = 0.57,
    colors = color.palette["duplication"],
    labels = "Type: Duplication",
  )
}

# breakends
bnd <- sv[sv$type == "BND",]
if(nrow(bnd) > 0){
  tracks <- tracks + BioCircosSNPTrack(
    'Breakends',
    as.character(bnd$contig),
    bnd$position_start,
    values = 0,
    minRadius = 0.33,
    maxRadius = 0.42,
    size = 1,
    shape = "rect",
    opacities = 0.7,
    colors = color.palette["breakend"],
    labels = "Type: Breakend"
  )
}

# loop to create backgrounds
for(i in c("inversion", "deletion", "duplication")){
  tracks <- tracks + BioCircosBackgroundTrack(
    paste0(i,"_background"),
    minRadius = radius, maxRadius = radius + 0.15,
    fillColors = "#F9F9F9",
    borderColors = "white",
    borderSize = 1
  )
  radius <- radius + 0.15
}

BioCircos(
  tracks,
  displayGenomeBorder = F,
  genome = genomeChr,
  chrPad = 0.03,
  genomeTicksDisplay = F,
  genomeLabelDy = 3,
  genomeLabelTextSize = 7,
  SNPMouseOverCircleSize = 2,
  SNPMouseOverTooltipsHtml03 = "<br/>Length: ",
  width = "100%",
  height = "1000px",
  elementId = "circosplot" 
)
```