---
title: "NAIBR SV - `r params$sample`"
include-in-header:
  - text: |
      <style>
      .BioCircosSNPTooltip {
        z-index: 10;
      }
      </style>
  - text: |
      <style>
      .BioCircosARCTooltip {
        z-index: 10;
      }
      </style>
params:
  faidx: 'genome.fasta.fai'
  bedpe: 'sample.bedpe'
  sample: 'sample'
  contigs: 'default'
---
`r format(Sys.time(), '🗓️ %d %B, %Y 🕔 %H:%M')`

```{r input_files}
infile <- params$bedpe
fai <- params$faidx
samplename <- params$sample
```

# General Stats
```{r read_input, results = 'asis'}
tryCatch(
  expr = {
    variants <- read.table(infile, header = T) 
    if (nrow(variants) == 0) {
      cat(paste0("There are no variants in the file ", "`", infile, "`"))
      knitr::knit_exit()
    }
  },
  error = function(e){ 
    cat(paste0("There are no variants in the file ", "`", infile, "`"))
    knitr::knit_exit()
  }
)
```

```{r package_imports}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("dplyr","tidyr","DT", "BioCircos")
```

```{r process_sv}
cleanbedpe <- function(x){
  y <- x[x$Chr1==x$Chr2, c(-3,-10)]
  y$Score <- round(y$Score, digits = 3)
  y$Length <- abs(y$Break2 - y$Break1)
  names(y)[1] <- "Chr1"
  return(y)
}

chimeric <- function(x){
  y <- x[x$Chr1!=x$Chr2, -10]
  y$Score <- round(y$Score, digits = 3)
  names(y)[1] <- "Chr1"
  return(y)
}

sv <- cleanbedpe(variants)
```

## 
```{r summary_stats}
summstats <- sv %>%
  group_by(Chr1, SV) %>%
  summarise(n = n())

sumtable <- summstats %>%
  group_by(SV) %>%
  summarise(count = sum(n))

nvar <- sum(sumtable$count)
```

```{r}
#| content: valuebox
#| title: "Variants"
list(
  color = ifelse(nvar > 0, "success", "warning"),
  value = scales::comma(nvar)
)
```

```{r}
#| content: valuebox
#| title: "Deletions"
ndel <- ifelse("deletion" %in% sumtable$SV, sumtable$count[which(sumtable$SV == "deletion")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ndel)
)
```

```{r}
#| content: valuebox
#| title: "Duplications"
ndup <- ifelse("duplication" %in% sumtable$SV, sumtable$count[which(sumtable$SV == "duplication")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ndup)
)
```

```{r}
#| content: valuebox
#| title: "Inversions"
ninv <- ifelse("inversion" %in% sumtable$SV, sumtable$count[which(sumtable$SV == "inversion")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ninv)
)
```

##
This report details the structural variants identified by [NAIBR](https://github.com/pontushojer/NAIBR)
[(publication)](https://doi.org/10.1093/bioinformatics/btx712). This tab (`General Stats`) shows overview
information. The `Per-Contig Plot` tab shows you an interactive
plot detailing all variants identified. The variants are given by:

Inversion
: A segment that broke off and reattached within the same chromosome, but in reverse orientation

Duplication
: A type of mutation that involves the production of one or more copies of a gene or region of a chromosome

Deletion
: A type of mutation that involves the loss of one or more nucleotides from a segment of DNA

## Various Stats Tables
::: {.card title="Variants by Count"}

This table details the count of the structural variants NAIBR detected.
```{r sv_count_table}
grpstats <- sv %>%
  group_by(Chr1, SV) %>%
  summarise(count = n(), total_bp  = sum(Length))

DT::datatable(
  grpstats %>% pivot_wider(id_cols = 1, names_from = SV, values_from = count),
  rownames = F,
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename , "_sv_naibr_count"))),
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::

###
::: {.card title="Variants by Length"}

This table details the total base pairs associated with the different types of structural variants. 
```{r sv_by_len_table}
DT::datatable(
  grpstats %>% pivot_wider(id_cols = 1, names_from = SV, values_from = total_bp),
  rownames = F,
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename , "_sv_naibr_count_bp"))),
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::


## Variants and Chimeras
::: {.card title="Non-Chimeric Variants"}

This table details the variants detected by NAIBR that appear on a single
contig/chromosome and passed the programs internal filtering. In other words, the
detected non-chimeric variants.

```{r all_variants_table}
DT::datatable(
  sv[,-10],
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_sv_naibr"))),
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::


###
::: {.card title="Chimeric Variants"}

This table shows any structural variants whose breakpoints span multiple contigs
("chimeric") and may require further assessment. These variants are omitted from plotting.

```{r chimeric_table}
DT::datatable(
  chimeric(variants),
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_sv_naibr_chimeric"))),
    scrollX = TRUE
  ),
  fillContainer = T
)
```
:::

# Per-Contig Plot
```{r filter_contigs}
fa.sizes <- read.table(fai, header = F) %>% arrange(desc(2))
colnames(fa.sizes) <- c("contig", "size")
plot_contigs <- params$contigs
if (all(plot_contigs == "default")){
  # make sure that the contigs with SV are the ones being plotted, not the others
  fa.sizes <- fa.sizes[fa.sizes$contig %in% unique(sv$Chr1), 1:2]
  # limit the data to only the 30 of the largest present contigs
  fa.sizes <- fa.sizes[1:min(nrow(fa.sizes), 30), ]
} else {
  fa.sizes <- fa.sizes[fa.sizes$contig %in% plot_contigs, 1:2]
}
# filter out variants that aren't on one of these contigs
sv <- sv[sv$Chr1 %in% fa.sizes$contig, ]
sv$Chr1 <- factor(sv$Chr1)
```
##
To the right is a circular plot to visualize the distribution of _putative_ structural variants
across (up to) 30 of the largest contigs, or whichever contigs were provided. If you are unfamiliar with this kind of visualization,
it's a circular representation of a linear genome. Each "wedge" is a different contig, from position 0 to 
the end of the contig, and is labelled by the contig name. Each internal (grey) ring is a plot
of observed structural variants for that contig. The legend shows which colors correspond to which type of variant. To save the plot,
you will need to take a screenshot. Very small variants may be difficult to see or
may not appear on the plot. Deletions are shown as points as they tend to be small
and clump together. If for some reason the plot isn't appearing, try refreshing
your browser.


##
::: {.card width=15% title="Legend"}
```{r}
calculate_luminance <- function(hex) {
  rgb_values <- col2rgb(hex) / 255
  luminance <- 0.2126 * rgb_values[1] + 0.7152 * rgb_values[2] + 0.0722 * rgb_values[3]
  return(luminance)
}

# Function to decide text color based on background luminance
get_text_color <- function(background_color) {
  luminance <- calculate_luminance(background_color)
  if (luminance > 0.5) {
    # Dark background -> Light text (black)
    return("black") 
  } else {
    # Light background -> Dark text (white)
    return("white") 
  }
}
color.palette <- c("deletion" = "#5a8c84", "duplication" = "#99278a", "inversion" = "#4a9fea")
DT::datatable(
  data.frame(Variant = c("deletion","duplication", "inversion")),
  rownames = F,
  fillContainer = T,
  colnames = "Variant Type",
  options = list(
    dom = 't',
    scrollX = F,
    scrollY = F,
    paging = F
  )
) %>% formatStyle(
  "Variant",
  backgroundColor = styleEqual(names(color.palette), color.palette),
  # Adjust text color dynamically
  color = styleEqual(names(color.palette), sapply(color.palette, get_text_color))
)
```
:::

###
```{r interactive_circos}
#| title: "Structural Variants (save by taking a screenshot)"

genomeChr <- fa.sizes$size
names(genomeChr) <- fa.sizes$contig
genomeChr <- as.list(genomeChr)
tracks <- BioCircosTracklist()
radius <- 0.3

# Add one track for each SV type
# inversions
inv <- sv[sv$SV == "inversion",]
if(nrow(inv) > 0){
  tracks <- tracks + BioCircosArcTrack(
    'Inversions',
    as.character(inv$Chr1),
    inv$Break1,
    inv$Break2,
    opacities = 0.7,
    values = inv$SplitMolecules,
    minRadius = 0.73,
    maxRadius = 0.87,
    colors = color.palette["inversion"],
    labels = "Type: Inversion",
  )
}

# deletions
del <- sv[sv$SV == "deletion",]
if(nrow(del) > 0){
  tracks <- tracks + BioCircosSNPTrack(
    'Deletions',
    as.character(del$Chr1),
    del$Break1,
    values = abs(del$Break2 - del$Break1),
    minRadius = 0.53,
    maxRadius = 0.67,
    size = 1,
    shape = "rect",
    opacities = 0.7,
    colors = color.palette["deletion"],
    labels = "Type: Deletion"
  )
}

# duplications
dup <- sv[sv$SV == "duplication",]
if(nrow(dup) > 0){
  tracks <- tracks + BioCircosArcTrack(
    'Duplications',
    as.character(dup$Chr1),
    dup$Break1,
    dup$Break2,
    values = dup$SplitMolecules,
    opacities = 0.7,
    minRadius = 0.33,
    maxRadius = 0.47,
    colors = color.palette["duplication"],
    labels = "Type: Duplication",
  )
}

# loop to create backgrounds
for(i in c("inversion", "deletion", "duplication")){
  tracks = tracks + BioCircosBackgroundTrack(
    paste0(i,"_background"),
    minRadius = radius, maxRadius = radius + 0.2,
    fillColors = "#F9F9F9",
    borderColors = "#C9C9C9",
    borderSize = 1
  )
  radius <- radius + 0.2
}

BioCircos(
  tracks,
  displayGenomeBorder = F,
  genome = genomeChr,
  chrPad = 0.03,
  genomeTicksDisplay = F,
  genomeLabelDy = 3,
  genomeLabelTextSize = 20,
  SNPMouseOverCircleSize = 2,
  SNPMouseOverTooltipsHtml03 = "<br/>Length: ",
  width = "100%",
  height = "1000px",
  elementId = "circosplot" 
)
```


# Supporting Info
##
::: {.card title="Interpreting NAIBR Output"}

NAIBR outputs a tab-delimited file with named columns (along with a VCF
and a reformatted bedpe file with extra information). The columns of the
bedpe file are deciphered as such:

```{r description_table}
knitr::kable(
  data.frame(
    "Column" = names(variants),
    "Descrption" = c(
      "Name of the contig where the variant starts",
      "Base-pair position in `Chr1` of the start of the variant",
      "Name of the contig where the variant ends",
      "Base-pair position in `Chr2` of the end of the variant",
      "Number of split molecules supporting variant",
      "Number of discordant reads supporting variant",
      "Orientation of variant relative to reference",
      "The haplotype of the variant",
      "log-likelihood score for variant",
      "`PASS` if passes internal NAIBR filter threshold, otherwise `FAIL`",
      "The type of variant (inversion, deletion, or duplication)"
    )
  ), col.names = c("Column Name", "Description")
)
```
:::