---
title: "NAIBR SV - `r params$sample`"
brand:
  logo: 
    medium: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_navbar.png
execute:
    echo: false
    warning: false
embed-resources: true
format: 
  dashboard:
    theme: materia
    scrolling: true
    expandable: false
    logo: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_navbar.png
    nav-buttons:
      - icon: book-half
        href: https://pdimens.github.io/harpy
      - icon: bug
        href: https://github.com/pdimens/harpy/issues/new/choose
      - icon: code-slash
        href: https://github.com/pdimens/harpy
params:
  faidx: 'genome.fasta.fai'
  bedpe: 'sample.bedpe'
  sample: 'sample'
  contigs: 'default'
---
`r format(Sys.time(), '🗓️ %d %B, %Y 🕔 %H:%M')`

```{r input_files}
infile <- params$bedpe
fai <- params$faidx
samplename <- params$sample
```

# General Stats
```{r read_input, results = 'asis'}
tryCatch(
  expr = {
    variants <- read.table(infile, header = T) 
    if (nrow(variants) == 0) {
      cat(paste0("There are no variants in the file ", "`", infile, "`"))
      knitr::knit_exit()
    }
  },
  error = function(e){ 
    cat(paste0("There are no variants in the file ", "`", infile, "`"))
    knitr::knit_exit()
  }
)
```

```{r package_imports}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("dplyr","tidyr", "ggplot2","DT", "BioCircos")
```

```{r process_sv}
cleanbedpe <- function(x){
  y <- x[x$Chr1==x$Chr2, c(-3,-10)]
  y$Score <- round(y$Score, digits = 3)
  y$Length <- abs(y$Break2 - y$Break1)
  names(y)[1] <- "Chr1"
  return(y)
}

chimeric <- function(x){
  y <- x[x$Chr1!=x$Chr2, -10]
  y$Score <- round(y$Score, digits = 3)
  names(y)[1] <- "Chr1"
  return(y)
}

sv <- cleanbedpe(variants)
```

## 
```{r summary_stats}
summstats <- sv %>%
  group_by(Chr1, SV) %>%
  summarise(n = n())

sumtable <- summstats %>%
  group_by(SV) %>%
  summarise(count = sum(n))

nvar <- sum(sumtable$count)
```

```{r}
#| content: valuebox
#| title: "Variants"
list(
  color = ifelse(nvar > 0, "success", "warning"),
  value = scales::comma(nvar)
)
```

```{r}
#| content: valuebox
#| title: "Deletions"
ndel <- ifelse("deletion" %in% sumtable$SV, sumtable$count[which(sumtable$SV == "deletion")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ndel)
)
```

```{r}
#| content: valuebox
#| title: "Duplications"
ndup <- ifelse("duplication" %in% sumtable$SV, sumtable$count[which(sumtable$SV == "duplication")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ndup)
)
```

```{r}
#| content: valuebox
#| title: "Inversions"
ninv <- ifelse("inversion" %in% sumtable$SV, sumtable$count[which(sumtable$SV == "inversion")], 0)
list(
  color = "#dfdfdf",
  value = scales::comma(ninv)
)
```

##
This report details the structural variants identified by [NAIBR](https://github.com/pontushojer/NAIBR)
[(publication)](https://doi.org/10.1093/bioinformatics/btx712). This tab (`General Stats`) shows overview
information. The `Per-Contig Plot` tab shows you an interactive
plot detailing all variants identified. The variants are given by:

Inversion
: A segment that broke off and reattached within the same chromosome, but in reverse orientation

Duplication
: A type of mutation that involves the production of one or more copies of a gene or region of a chromosome

Deletion
: A type of mutation that involves the loss of one or more nucleotides from a segment of DNA

## Various Stats Tables
::: {.card title="Variants by count"}

This table details the count of the structural variants NAIBR detected.
```{r sv_count_table}
grpstats <- sv %>%
  group_by(Chr1, SV) %>%
  summarise(count = n(), total_bp  = sum(Length))

DT::datatable(
  grpstats %>% pivot_wider(id_cols = 1, names_from = SV, values_from = count),
  rownames = F,
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename , "_sv_naibr_count"))),
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::

###
::: {.card title="Variants by Length"}

This table details the total base pairs associated with the different types of structural variants. 
```{r sv_by_len_table}
DT::datatable(
  grpstats %>% pivot_wider(id_cols = 1, names_from = SV, values_from = total_bp),
  rownames = F,
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename , "_sv_naibr_count_bp"))),
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::


## Variants and Chimeras
::: {.card title="Non-chimeric variants"}

This table details the variants detected by NAIBR that appear on a single
contig/chromosome and passed the programs internal filtering. In other words, the
detected non-chimeric variants.

```{r all_variants_table}
DT::datatable(
  sv[,-10],
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_sv_naibr"))),
    scrollX = T,
    paging = T
  ),
  fillContainer = T
)
```
:::


###
::: {.card title="Chimeric Variants"}

This table shows any structural variants whose breakpoints span multiple contigs
("chimeric") and may require further assessment. These variants are omitted from plotting.

```{r chimeric_table}
DT::datatable(
  chimeric(variants),
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = paste0(samplename ,"_sv_naibr_chimeric"))),
    scrollX = TRUE
  ),
  fillContainer = T
)
```
:::

# Per-Contig Plot
```{r filter_contigs}
fa.sizes <- read.table(fai, header = F) %>% arrange(desc(2))
colnames(fa.sizes) <- c("contig", "size")
plot_contigs <- params$contigs
if (all(plot_contigs == "default")){
  # make sure that the contigs with SV are the ones being plotted, not the others
  fa.sizes <- fa.sizes[fa.sizes$contig %in% unique(sv$Chr1), 1:2]
  # limit the data to only the 30 of the largest present contigs
  fa.sizes <- fa.sizes[1:min(nrow(fa.sizes), 30), ]
} else {
  fa.sizes <- fa.sizes[fa.sizes$contig %in% plot_contigs, 1:2]
}
# filter out variants that aren't on one of these contigs
sv <- sv[sv$Chr1 %in% fa.sizes$contig, ]
sv$Chr1 <- factor(sv$Chr1)
```
##
To the right is a circular plot to visualize the distribution of _putative_ structural variants
across (up to) 30 of the largest contigs, or whichever contigs were provided. If you are unfamiliar with this kind of visualization,
it's a circular version of a linear genome. Each "wedge" is a different contig, from position 0 to 
the end of the contig, and is labelled by the contig name. Each internal (grey) ring is a plot
of observed structural variants for that contig. The legend shows which colors correspond to which type of variant. To save the plot,
you will need to take a screenshot. Very small variants may be difficult to see or
may not appear on the plot. Deletions are shown as points as they tend to be small
and clump together.


##
::: {.card width=20% title="Legend"}
```{r sv_legend_plot, dev="png"}
color.palette <- c("deletion" = "#5a8c84", "duplication" = "#99278a", "inversion" = "#4a9fea")
data.frame(x = 1, y = 1:3, colour = c("inversion", "deletion","duplication")) %>% 
ggplot(aes(x, y, fill = colour))+
  geom_point(alpha=0, shape = 22, color = "white")+ # completely transparent rectangular point 
  scale_fill_manual(values=color.palette, drop=FALSE) +
  guides(fill = guide_legend(override.aes = list(alpha=1, size = 25)))+ # showing the point in the legend
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(0.3, 0.5),
        legend.text = element_text(size = 28),
        legend.title=element_text(size=28),
        legend.key = element_rect(fill='NA'),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), 
        plot.background = element_rect(fill='transparent', color=NA),
        legend.background = element_rect(fill='transparent'),
        legend.box.background = element_rect(fill='transparent', color = "transparent"), 
  ) +
  labs(fill="Variant Type")
```
:::

###
```{r interactive_circos}
#| title: "Structural Variants (save by taking a screenshot)"

genomeChr <- fa.sizes$size
names(genomeChr) <- fa.sizes$contig
genomeChr <- as.list(genomeChr)
tracks <- BioCircosTracklist()
radius <- 0.3

# Add one track for each SV type
# inversions
inv <- sv[sv$SV == "inversion",]
if(nrow(inv) > 0){
  tracks <- tracks + BioCircosArcTrack(
    'Inversions',
    as.character(inv$Chr1),
    inv$Break1,
    inv$Break2,
    opacities = 0.7,
    values = inv$SplitMolecules,
    minRadius = 0.73,
    maxRadius = 0.87,
    colors = color.palette["inversion"],
    labels = "Type: Inversion",
  )
}

# deletions
del <- sv[sv$SV == "deletion",]
if(nrow(del) > 0){
  tracks <- tracks + BioCircosSNPTrack(
    'Deletions',
    as.character(del$Chr1),
    del$Break1,
    values = abs(del$Break2 - del$Break1),
    minRadius = 0.53,
    maxRadius = 0.67,
    size = 1,
    shape = "rect",
    opacities = 0.7,
    colors = color.palette["deletion"],
    labels = "Type: Deletion"
  )
}

# duplications
dup <- sv[sv$SV == "duplication",]
if(nrow(dup) > 0){
  tracks <- tracks + BioCircosArcTrack(
    'Duplications',
    as.character(dup$Chr1),
    dup$Break1,
    dup$Break2,
    values = dup$SplitMolecules,
    opacities = 0.7,
    minRadius = 0.33,
    maxRadius = 0.47,
    colors = color.palette["duplication"],
    labels = "Type: Duplication",
  )
}

# loop to create backgrounds
for(i in c("inversion", "deletion", "duplication")){
  tracks = tracks + BioCircosBackgroundTrack(
    paste0(i,"_background"),
    minRadius = radius, maxRadius = radius + 0.2,
    fillColors = "#F9F9F9",
    borderColors = "white",
    borderSize = 1
  )
  radius <- radius + 0.2
}

BioCircos(
  tracks,
  displayGenomeBorder = F,
  genome = genomeChr,
  chrPad = 0.03,
  genomeTicksDisplay = F,
  genomeLabelDy = 3,
  genomeLabelTextSize = 20,
  SNPMouseOverCircleSize = 2,
  SNPMouseOverTooltipsHtml03 = "<br/>Length: ",
  width = "100%",
  height = "1000px",
  elementId = "circosplot" 
)
```


# Supporting Info
##
::: {.card title="Interpreting NAIBR Output"}

NAIBR outputs a tab-delimited file with named columns (along with a VCF
and a reformatted bedpe file with extra information). The columns of the
bedpe file are deciphered as such:

```{r description_table}
knitr::kable(
  data.frame(
    "Column" = names(variants),
    "Descrption" = c(
      "Name of the contig where the variant starts",
      "Base-pair position in `Chr1` of the start of the variant",
      "Name of the contig where the variant ends",
      "Base-pair position in `Chr2` of the end of the variant",
      "Number of split molecules supporting variant",
      "Number of discordant reads supporting variant",
      "Orientation of variant relative to reference",
      "The haplotype of the variant",
      "log-likelihood score for variant",
      "`PASS` if passes internal NAIBR filter threshold, otherwise `FAIL`",
      "The type of variant (inversion, deletion, or duplication)"
    )
  ), col.names = c("Column Name", "Description")
)
```
:::