---
title: "FASTQ Haplotag Format Check"
date: "`r format(Sys.time(), '%d %b, %Y at %H:%M')`"
brand:
  logo: 
    medium: logo.png
execute:
    echo: false
    warning: false
embed-resources: true
format: 
  dashboard:
    theme:
      light: materia
      dark: slate
    scrolling: true
    logo: https://raw.githubusercontent.com/pdimens/harpy/docs/static/logo_report.png
    nav-buttons:
      - icon: book-half
        href: https://pdimens.github.io/harpy
      - icon: bug
        href: https://github.com/pdimens/harpy/issues/new/choose
      - icon: code-slash
        href: https://github.com/pdimens/harpy
params:
  infile: 'test.txt'
---

```{r}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("dplyr","tidyr","DT","highcharter","scales")
```

```{r}
#infile <- "~/Documents/harpy/Preflight/fastq/filecheck.fastq.tsv"
infile <- params$infile

data <- read.table(infile, header = T)
attention_df <- data %>% 
  mutate(allmissing = (reads == noBX)) %>% 
  select(4,5,7) %>% 
  rowwise() %>% 
  summarise(count = sum(c(badBX,badSamSpec, allmissing)))
attention <- sum(attention_df$count > 0)
bxnotlast <- sum(data$bxNotLast > 0)
```
# File Stats

<h1> Preflight checks for FASTQ files </h1>

This report reflects the FASTQ files provided. Harpy has processed
the files to identify [formatting issues](#details) that may require your attention.
Below are a plot and table of the results of the preflight checks.

## row
```{r}
#| content: valuebox
#| title: "FASTQ Files"
list(
  icon = "files",
  color = "#C0C0C0",
  value = scales::comma(nrow(data))
)
```

```{r}
#| content: valuebox
#| title: "Files with Issues"
list(
  icon = ifelse(attention > 0, "exclamation-triangle", "check2-square"),
  color = ifelse(attention > 0, "warning", "success"),
  value = scales::comma(attention)
)
```

```{r}
#| content: valuebox
#| title: "BX:Z: tag not last"
list(
  icon = ifelse(attention > 0, "exclamation-triangle", "check2-square"),
  color = ifelse(bxnotlast > 0, "#feda75", "success"),
  value = scales::comma(attention)
)
```

## row

| data column | name in plot        | pass condition                                                             | fail condition                                              |
| :---------- | :------------------ | :------------------------------------------------------------------------- | :---------------------------------------------------------- |
| badBX       | AxxCxxBxxDxx format | **all** reads with BX:Z: tag have properly formatted AxxCxxBxxDxx barcodes | **any** BX:Z: barcodes have incorrect format                |
| badSamSpec  | follows SAM spec    | **all** reads have proper TAG:TYPE:VALUE comments                          | **any** reads have incorrectly formatted comments           |
| bxNotLast   | BX:Z: last comment  | **all** reads have BX:Z: as final comment                                  | **at least 1 read** doesn’t have BX:Z: tag as final comment |
| noBX        | BX:Z: tag           | **any** BX:Z: tags present                                                 | **all** reads lack BX:Z: tag                                |

## File Stats
```{r status_df}
data2 <- data %>% 
  mutate(
    noBX = ifelse(noBX == reads, 0, 1),
    badBX = ifelse(badBX > 0, 0, 1),
    badSamSpec = ifelse(badSamSpec > 0, 0, 1),
    bxNotLast = ifelse(bxNotLast > 0, 0, 1)
    ) %>% 
  select(-2)
  m <- as.matrix(data2[,-1]) * 1
  colnames(m) <- c("AxxCxxBxxDxx format", "follows SAM Spec", "BX:Z: last comment", "BX:Z: tag")
  rownames(m) <- data$file
```

```{r readsper}
#| title: "File Preflight Checks (interactive)"
hchart(m) |>
  hc_colorAxis(stops = list(c(0, "#feda75"), c(1, "#4faad1")), tickAmount = 2, min = 0, max = 1) |>
  hc_title(text = "FASTQ File Checks") |>
  hc_xAxis(labels = list(style = list(fontSize = '16px')) )|>
  hc_yAxis(labels = list(style = list(fontSize = '16px')) )|>
  hc_plotOptions(heatmap = list(animation = FALSE, borderWidth = 1, borderColor = "white", states = list(inactive = list(animation = FALSE), normal = list(animation = FALSE)))) |>
  hc_legend(enabled = F) |>
  hc_tooltip(
    animation = FALSE,
    formatter = JS("function () {return '<b>' + this.series.xAxis.categories[this.point.x] + '</b><br>' + this.series.yAxis.categories[this.point.y] + '<br/>QC: <b>' + (this.point.value<1 ? 'FAIL</b>' : 'PASS</b>');}") 
  ) |>
  hc_exporting(enabled = T, filename = "fastq.preflight",
    buttons = list(contextButton = list(menuItems = c("downloadPNG", "downloadJPEG", "downloadPDF", "downloadSVG")))
  )
```

```{r dtable}
#| title: "File Preflight Checks (as a table)"
DT::datatable(
  data,
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  options = list(
    dom = 'Brtip',
    buttons = list(list(extend = "csv",filename = "fastq_checks")),
    scrollX = T
  ),
  fillContainer = T
)
```

# Details
## interpret
::: {.card title="Interpreting the Data"}

The `harpy preflight fastq ...` command created a `filecheck.fastq.tsv` file in the specified
output directory that summarizes the results that are included in this report.
This file contains a tab-delimited table with the columns: `file`,	`reads`, `noBX`,	`badBX`, and	`badSamSpec`. These columns are defined by:

file
: The name of the FASTQ file

reads
: The total number of reads in the file

noBX
: The number of reads that do not have a `BX:Z:` tag in the read header

- severity: <strong style='color:#c0c0c0;'>minimal</strong> and likely won't cause issues unless no reads have `BX:Z:` tags
- if you expect all of your reads should have `BX:Z:` tags, then further investigation is necessary

badBX
: The haplotag barcode in the `BX:Z:` comment does not adhere to the proper `AxxCxxBxxDxx` format

- severity: <strong style='color:red;'>critical</strong> and will cause downstream issues and errors

badSamSpec
: The comments in the read header after the read ID do not conform to the `TAG:TYPE:VALUE` [SAM specification](https://samtools.github.io/hts-specs/SAMv1.pdf)

- severity: <strong style='color:red;'>critical</strong> and will cause downstream issues and errors

bxNotLast
: The `BX:Z:` tag in the FASTQ header is not the last comment

- severity: <strong style='color:orange;'>conditional</strong> and only relevant for LEVIATHAN variant calling 
- can be ignored if not intending to call structural variants with LEVIATHAN
- if intending to use LEVIATHAN, the `BX:Z:` tag must be the last comment in a read header
- `harpy align -m bwa` automatically moves the `BX:Z:` tag to the end of the alignment record
:::

### col
::: {.card title="Proper Read Headers"}

SAM Specification
: An example of a proper FASTQ read header is like the one below, where all comments
following the initial read ID (`@A00814...`) have the proper [SAM spec](https://samtools.github.io/hts-specs/SAMv1.pdf)
`TAG:TYPE:VALUE` format and among them is a `BX:Z:` tag followed by a `AxxCxxBxxDxx`
formatted haplotag barcode. 

Comments
: The comments must be **TAB separated** because the `:Z:` tag type allows a 
whitespace character for its value. 

Example Header
: The read header below contains the read ID `@A00814...`,
the `BX:Z:` haplotag barcode tag, and two more comments `RX:Z:` and `QX:Z:` that both
adhere to the SAM specification. If using LEVIATHAN to call structural variants, the `BX:Z:` tag
must be the last comment in the read header.
```
@A00814:267:HTMH3DRXX:2:1101:4580:1000	BX:Z:A02C01B11D46	RX:Z:GAAACGACCAACA+CGAACACGTTAGC	QX:Z:F,FFFFFFFFFFF+FF,FFF:FFFFFF
```
:::