---
title: "FASTQ Haplotag Format Check"
params:
  infile: 'test.tsv'
  platform: 'haplotagging'
---
`r format(Sys.time(), '🗓️ %d %B, %Y 🕔 %H:%M')`

```{r}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need, repos = "https://cloud.r-project.org/")
        lapply(need,require,character.only=TRUE)
    }
}
using("dplyr","tidyr","DT","scales")
```

```{r}
infile <- params$infile

data <- read.table(infile, header = T)
attention_df <- data %>% 
  mutate(allmissing = (reads == noBX)) %>% 
  select(4,5,7) %>% 
  rowwise() %>% 
  summarise(count = sum(c(badBX,badSamSpec, allmissing)))
attention <- sum(attention_df$count > 0)
bxnotlast <- sum(data$bxNotLast > 0)
```
# File Stats
<h1> Preflight checks for FASTQ files </h1>

This report reflects the FASTQ files Harpy processed to identify obvious
[formatting issues](#details) that may require your attention.

## row
```{r}
#| content: valuebox
#| title: "FASTQ Files"
list(
  icon = "files",
  color = "light",
  value = scales::comma(nrow(data))
)
```

```{r}
#| content: valuebox
#| title: "Files with Issues"
list(
  icon = ifelse(attention > 0, "exclamation-triangle", "check2-square"),
  color = ifelse(attention > 0, "warning", "success"),
  value = scales::comma(attention)
)
```

```{r}
#| content: valuebox
#| title: "BX Tag Not Last"
list(
  icon = ifelse(attention > 0, "exclamation-triangle", "check2-square"),
  color = ifelse(bxnotlast > 0, "warning", "success"),
  value = scales::comma(bxnotlast)
)
```

## desc

| column           | 🟦 pass condition 🟦                                                        | 🟨 fail condition 🟨                                     |
| :--------------- | :---------------------------------------------------------------------------- | :--------------------------------------------------------- |
| **Format** | **all** reads with `BX:Z` tag have properly formatted `{r platform}` barcodes | **any** BX:Z barcodes have incorrect `{r platform}` format                |
| **SAM Spec**     | **all** reads have proper `TAG:TYPE:VALUE` comments                           | **any** reads have incorrectly formatted comments          |
| **BX:Z last**    | **all** reads have `BX:Z` as final comment                                    | **at least 1 read** doesn’t have BX:Z tag as final comment |
| **BX:Z tag**     | **any** `BX:Z` tags present                                                   | **all** reads lack BX:Z tag                                |

## File Stats
```{r resultstable}
#| title: "File Stats"

datatable(
  data,
  escape = F,
  rownames = F,
  filter = "top",
  extensions = 'Buttons',
  fillContainer = T,
  colnames = c("File", "Reads", "AxxCxxBxxDxx", "SAM Spec", "BX:Z last", "BX:Z tag"),
  options = list(
    paging = F,
    scrollX = F,
    dom = 'Brtp',
    buttons = list(list(extend = "csv",filename = "fastq_checks"))
  )
) %>%
formatStyle(
  columns = 3:ncol(data),
  # Blue for values <= 0, yellow for values > 0
  backgroundColor = styleInterval(0, c('#4faad1', '#feda75')) 
) %>%
formatStyle(
  columns = 2,
  # yellow for values == 0, blue for values > 0
  backgroundColor = styleInterval(0, c('#feda75', '#4faad1')) 
)     
```


# Details
## interpret
::: {.card title="Interpreting the Data"}

The `harpy validate fastq ...` command created a `filecheck.fastq.tsv` file in the specified
output directory that summarizes the results that are included in this report.
This file contains a tab-delimited table with the columns: `file`,	`reads`, `noBX`,	`badBX`, and	`badSamSpec`.
These columns are defined below and the severities of their issues are given as colored emoji:

- ⬜ is minor
- 🔶 is moderate
- 🛑 is critical

file
: The name of the FASTQ file.

reads
: The total number of reads in the file.

noBX ⬜
: The number of reads that do not have a `BX:Z` tag in the read header.
If you expect all or some of your reads should have `BX:Z` tags, then further investigation is necessary.

badBX 🛑
: The haplotag barcode in the `BX:Z` comment does not adhere to the proper `AxxCxxBxxDxx` format.

badSamSpec 🛑
: The comments in the read header after the read ID do not conform to the `TAG:TYPE:VALUE` [SAM specification](https://samtools.github.io/hts-specs/SAMv1.pdf).

bxNotLast 🔶
: The `BX:Z:` tag in the FASTQ header is not the last comment. 
Only relevant for LEVIATHAN variant calling and can be ignored if not intending to call
structural variants with LEVIATHAN,otherwise the `BX:Z` tag must be the last comment in a read
header. `harpy align` will automatically move the `BX:Z` tag to the end of the alignment record.
:::

### col
::: {.card title="Proper Read Headers"}

SAM Specification
: An example of a proper FASTQ read header is like the one below, where all comments
following the initial read ID (`@A00814...`) have the proper [SAM spec](https://samtools.github.io/hts-specs/SAMv1.pdf)
`TAG:TYPE:VALUE` format and among them is a `BX:Z` tag followed by a `AxxCxxBxxDxx`
formatted haplotag barcode. 

Comments
: The comments must be **TAB separated** because the `:Z` tag type allows a 
whitespace character for its value. 

Example Header
: The read header below contains the read ID `@A00814...`,
the `BX:Z` haplotag barcode tag, and two more comments `RX:Z` and `QX:Z` that both
adhere to the SAM specification. If using LEVIATHAN to call structural variants, the `BX:Z` tag
must be the last comment in the read header.
```
@A00814:267:HTMH3DRXX:2:1101:4580:1000	BX:Z:A02C01B11D46	RX:Z:GAAACGACCAACA+CGAACACGTTAGC	QX:Z:F,FFFFFFFFFFF+FF,FFF:FFFFFF
```

:::